---
title: "DPA"
output:
  html_document: default
  pdf_document: default
date: "2025-05-27"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Divergence Point Analysis on Eye-tracking Data

This R script generates time-course plots of mean fixation proportion, and performs a statistical analysis to determine divergence points between fixations on different objects. It is adapted from the R script in the paper "Analyzing data from the psycholinguistic visual‑world paradigm: Comparison of different analysis methods", Ito & Knoeferle (2023).

The columns in the original eye-tracking data set are:

-   TOI: trial of interest. There are 48 trials in Version A and 48 trials in Version B.

-   Participant: the code for a participant. There are 32 participants in total; half the participants were assigned to Version A trials and half were assigned to Version B trials.

-   Timeline: the version of the experiment

-   Bin: the number of the time bin for one participant and trial

-   Bin_duration: the duration of each time bin

-   Competitor: binomially coded fixation data - 1=fixation on competitor object in the current time bin; 0=fixation elsewhere

-   Distractor: binomially coded fixation data - 1=fixation on distractor object in the current time bin; 0=fixation elsewhere

-   Target: binomially coded fixation data - 1=fixation on target object in the current time bin; 0=fixation elsewhere

-   TimeMS: the time bin converted to time in milliseconds

-   Animacy_C: the animacy of the competitor noun

-   Animacy_D: the animacy of the distractor noun

-   Animacy_T: the animacy of the target noun

-   Determiner: the type of determiner in the trial (Definite or Possessive)

-   N_Gender_C: the lexical gender of the competitor noun (Masculine or Feminine)

-   N_Gender_D: the lexical gender of the distractor noun (Masculine or Feminine)

-   N_Gender_T: the lexical gender of the target noun (Masculine or Feminine)

-   Phonol_Overlap_C: whether the competitor noun has a minimal pair that only differs in the word-final vowel and lexical gender

-   Phonol_Overlap_D: whether the distractor noun has a minimal pair that only differs in the word-final vowel and lexical gender

-   Phonol_Overlap_T: whether the target noun has a minimal pair that only differs in the word-final vowel and lexical gender

-   ArticleOnset: the onset time (in milliseconds) of the determiner in the audio cue of the current trial

-   NounOnset: the onset time (in milliseconds) of the noun in the audio cue of the current trial

-   EndingOnset: the onset time (in milliseconds) of the noun's final vowel in the audio cue of the current trial

-   EndingOffset: the offset time (in milliseconds) of the noun's final vowel in the audio cue of the current trial

-   Remaining: the remaining time after the offset of the audio cue of the current trial

-   SpeakerType: the bilingual status of participant - 1 (Adult Spanish Immigrant) or 2 (Heritage Spanish Speaker)

-   Sex: the sex of the participant (Femenino or Masculino)

-   Age: the age of the participant

-   Position: the position of the object in fixation (top left, bottom centre, top right)

-   Trackloss: whether the fixation is in blank space or off screen, or there is a blink in the current time bin (TRUE or FALSE)

-   TimeMS_adjusted: TimeMS zeroed at ArticleOnset

-   ArticleOnset_adjusted: ArticleOnset zeroed at ArticleOnset (thus always 0)

-   NounOnset_adjusted: NounOnset zeroed at ArticleOnset

-   EndingOnset_adjusted: EndingOnset zeroed at ArticleOnset

-   EndingOffset_adjusted: EndingOffset zeroed at ArticleOnset

-   Remaining_adjusted: Remaining zeroed at ArticleOnset

### 1. Data preparation

```{r eval=T, echo=T, message=F}
rm(list = ls()) # Clear the workspace
options(scipen = 999) # Turn off scientific notation
```

##### Packages

```{r eval=T, message=F}
require(Rmisc)
require(tidyverse)
require(ggplot2)
require(boot)
require(dplyr)
require(lme4)
require(purrr)
```

##### Package versions

```{r eval=T, echo=F}
installed.packages()[names(sessionInfo()$otherPkgs), "Version"]
```

##### Read the time-course output data of the eye-tracker from file

```{r eval=T, message=F}
fix.dat = read_delim("data-2025-03-12/df_cleaned.csv", delim=',')
fix.dat = fix.dat %>% mutate_at(vars(Timeline), as.factor) # TODO not sure why this is changed into a factor now
# all the trackloss bins have been removed - in any given row, there is a fixation in one of the 3 AreaFixated 
```

##### Inspect the raw data

```{r eval=T, message=F}
summary(fix.dat)
head(fix.dat)
```

##### Examine which participants did which trials

```{r eval=T}
pt_matrix <- fix.dat %>%
  distinct(Participant, TOI) %>%
  mutate(did_trial = 1) %>%
  tidyr::pivot_wider(
    names_from = TOI,
    values_from = did_trial,
    values_fill = 0
  )

pt_matrix
```

##### Fill in the missing time points and add a trackloss column

```{r eval=T, echo=T, message=F}
# Step 1: Create complete time series based on the window of interest
# for all the possible time bins, use min(fix.dat$TimeMS_adjusted) and max(fix.dat$TimeMS_adjusted)

WINDOW_START <- min(fix.dat$TimeMS_adjusted)
WINDOW_END <- max(fix.dat$TimeMS_adjusted)
all_times <- seq(
  WINDOW_START,
  WINDOW_END,
  by = 10
)

# Step 2: record the participant-trial pairs that actually occurred in the raw data
observed_combos <- fix.dat %>%
  distinct(Participant, TOI)

# Step 3: fill in the missing time points for those participant-trial combos
full_template <- observed_combos %>%
  tidyr::expand(
    tidyr::nesting(Participant, TOI),
    TimeMS_adjusted = all_times
  )

# Step 4: left-join raw data onto the template
fix.dat.full <- full_template %>%
  left_join(fix.dat, by = c("Participant", "TOI", "TimeMS_adjusted"))
```

##### THE FOLLOWING IS BAD BECAUSE WE DON'T WANT TO FILL IN MISSING TIME POINTS FOR TRIALS A PARTICIPANT DID NOT DO {style="color: red"}

```{r eval=F, message=F}
# Step 2: Make sure fix.dat$TimeMS_adjusted is numeric
#fix.dat$TimeMS_adjusted <- as.numeric(fix.dat$TimeMS_adjusted)

# Create every Participant x Trial combination
participants <- unique(fix.dat$Participant)
TOIs <- unique(fix.dat$TOI)

# Build a table with all 32 × 96 combinations
all_combos <- fix.dat %>%
  distinct(Participant, TOI)

# Expand to include every time point for every combination
full_template <- all_combos %>%
  tidyr::expand(
    tidyr::nesting(Participant, TOI),
    TimeMS_adjusted = all_times
  )
```

##### Build a unique trial info table

```{r eval=T, message=F}
# TODO could add other trial variables as needed - if we are comparing by those variables
trial_info <- fix.dat %>%
  filter(!is.na(Determiner)) %>%
  group_by(TOI) %>%
  summarise(
    Determiner = unique(Determiner),
    # TODO don't think SpeakerType should be included in trial_info - it's not a property of a trial
    # SpeakerType = first(SpeakerType),
    NounOnset_adjusted = unique(NounOnset_adjusted),
    EndingOffset_adjusted = unique(EndingOffset_adjusted),
    .groups = "drop"
  )
```

##### Fill in missing information

```{r eval=T, message=F}
# Merge it in and fill NAs
fix.dat.full <- fix.dat.full %>%
  left_join(trial_info, by = "TOI", suffix = c("", ".trial")) %>%
  mutate(
    # fill AOI and Trackloss columns
    Trackloss  = ifelse(is.na(Trackloss), 1, Trackloss),
    Target     = ifelse(is.na(Target), 0, Target),
    Competitor = ifelse(is.na(Competitor), 0, Competitor),
    Distractor = ifelse(is.na(Distractor), 0, Distractor),
    
    # fill Determiner and other trial-level values
    Determiner = coalesce(Determiner, Determiner.trial),
    
    # TODO not sure why we can't fill in SpeakerType: Error in `mutate()`:
# ℹ In argument: `SpeakerType = coalesce(SpeakerType,
#   SpeakerType.trial)`.
# Caused by error:
# ! object 'SpeakerType.trial' not found

    # SpeakerType = coalesce(SpeakerType, SpeakerType.trial),
    NounOnset_adjusted = coalesce(NounOnset_adjusted, NounOnset_adjusted.trial),
    EndingOffset_adjusted = coalesce(EndingOffset_adjusted, EndingOffset_adjusted.trial)
  ) %>%
  select(-any_of(c("Determiner.trial", 
                   # "SpeakerType.trial",
                   "NounOnset_adjusted.trial", 
                   "EndingOffset_adjusted.trial")))
```

#### Sanity checks

```{r eval=T, message=F}
# Check that everyone now has 171 bins
fix.dat.full %>%
  group_by(Participant, TOI) %>%
  summarise(n_bins = n()) %>%
  summarise(
    n_missing = sum(n_bins != 171),
    mean_bins = mean(n_bins)
  )
# n_missing = 0
# mean_bins = 171

# Check the time range
range(fix.dat.full$TimeMS_adjusted)
# [1] -2670  4750

# Check a participant-trial pair where timepoints were formerly missing
fix.dat.full %>%
  filter(Participant == "TA106", TOI == "Trial01-def-campesina_0") %>%
  summarise(min_time = min(TimeMS_adjusted), max_time = max(TimeMS_adjusted))

# Check a participant-trial pair that's not supposed to be there (to make sure we didn't invent anything)
fix.dat.full %>%
  filter(Participant == "TA105", TOI == "Trial01-pos-campesina_0") %>%
  summarise(min_time = min(TimeMS_adjusted), max_time = max(TimeMS_adjusted))
# Should return Inf and -Inf

# Save this complete dataset
write_csv(fix.dat.full, "data/fix_dat_filled.csv")
# NOTE: there seems to be some extra rows at the end of each trial? Should be fine since we're dropping the rows outside our pre-defined time-window anyway

# Old implementation...
# Step 3: Merge original data with complete time series
# fix.dat <- fix.dat %>%
#   group_by(Participant, TOI) %>%
#   complete(TimeMS_adjusted = all_times) %>% # TODO do we want a full sequence of the time specific to the participant-trial group, or do we want to use the global min and max time?
#   # use full_seq(TimeMS_adjusted, 10) for local time range
#   mutate(Trackloss = ifelse(is.na(Trackloss), 1, 0)) %>%
#   mutate(Bin = row_number()) %>%
#   mutate(
#     Competitor = ifelse(Trackloss == 1, 0, Competitor),
#     Distractor = ifelse(Trackloss == 1, 0, Distractor),
#     Target = ifelse(Trackloss == 1, 0, Target)
#     ) %>%
#   ungroup()
# 
# # Step 4: Fill in the missing bin numbers
# bin_width <- 10  # or whatever your real bin size is
# 
# fix.dat <- fix.dat %>%
#   group_by(Participant, TOI) %>%
#   mutate(Bin = floor(TimeMS_adjusted/10+268)) %>%
#   ungroup()
```

##### Change `TOI` and `Participant` to factor

```{r eval=T, echo=T, message=F}
fix.dat.full = fix.dat.full %>% mutate_at(vars(TOI, Participant), as.factor)
TOIs <- sort(as.character(unique(fix.dat$TOI)))
participants <- as.character(unique(fix.dat$Participant))

# count how many unique participant–trial combinations exist in your dataset.
# it should be 96/2=48 trials * 32 participants = 1536, it seems like 4 combinations are missing because observed_combos has 1532 rows
```

##### Sanity check plot (step function plot for one participant and TOI)

```{r eval=T}
df_plot <- fix.dat.full %>%
  filter(Participant == "TA105", TOI == "Trial02-def-perro_0") %>%
  select(TimeMS_adjusted, Target, Competitor, Distractor) %>%
  pivot_longer(cols = c(Target, Competitor, Distractor),
               names_to = "AOI", values_to = "Fixation")

library(ggplot2)

vline_df <- tibble::tibble(
  time = c(108, 551),
  event = c("Noun onset", "Ending offset")
)

ggplot(df_plot, aes(x = TimeMS_adjusted, y = Fixation, color = AOI)) +
  geom_step(size = 0.8) +
  geom_vline(
    data = vline_df,
    aes(xintercept=time, linetype=event),
    color="black",
    linewidth=0.8
  )
  labs(title = "Binary Fixations by AOI",
       subtitle = "Participant TA105 – Trial02-def-perro_0",
       x = "Time (ms)",
       y = "Fixated (0/1)") +
  theme_minimal()

```

##### More sanity plotting

```{r eval=T}
# Set a trial and participant to inspect
trial_id <- "Trial01-def-campesina_0"
participant_id <- "TA105"

plot_df <- fix.dat.full %>%
  filter(TOI == trial_id, Participant == participant_id) %>%
  select(TimeMS_adjusted, Target, Competitor, Distractor) %>%
  pivot_longer(cols = c(Target, Competitor, Distractor),
               names_to = "AOI", values_to = "FixP")

ggplot(plot_df, aes(x = TimeMS_adjusted, y = FixP, color = AOI)) +
  geom_step(size = 1) +
  labs(title = paste("Step Function of Fixation Proportion"),
       subtitle = paste(participant_id, "–", trial_id),
       x = "Time (ms)", y = "Fixation Proportion") +
  scale_x_continuous(limits = c(-300, 1400)) +
  scale_y_continuous(limits = c(0, 1)) +
  theme_minimal()
```

##### Average noun onset and ending offset because our window feels quite small

```{r eval=F}
fix.dat.full %>%
  summarise(avg_onset = mean(unique(NounOnset_adjusted), na.rm = TRUE))
fix.dat.full %>%
  summarise(avg_offset = mean(unique(EndingOffset_adjusted), na.rm = TRUE))
```

# TODO don't forget to change fix.dat to fix.dat.full from this point onwards

# NOTE this is probably a past attempt at normalization

##### IGNORE: Re-bin the fixation data and store it in a new dataframe {style="color: red"}

```{r eval=F, echo=T, message=F}
fix.rebinned <- data.frame()

for (trial_id in TOIs[3:3]) {
  
  # Get scalar onset/offset for this trial
  noun_onset <- fix.dat %>%
    filter(TOI == trial_id, !is.na(NounOnset_adjusted)) %>%
    pull(NounOnset_adjusted) %>%
    unique()

  ending_offset <- fix.dat %>%
    filter(TOI == trial_id, !is.na(EndingOffset_adjusted)) %>%
    pull(EndingOffset_adjusted) %>%
    unique()
  
  # duration of each pseudo-bin
  pbin_dur <- (ending_offset - noun_onset) / 100  
  
  for (participant in participants[1:1]) {
    
    # Subset relevant rows for this trial × participant
    df_trial <- fix.dat[fix.dat$TOI == trial_id & 
                          fix.dat$Participant == participant &
                          fix.dat$TimeMS_adjusted >= noun_onset + 200 &
                          fix.dat$TimeMS_adjusted <= ending_offset + 200, ]
    
    for (i in 1:100) {
      start_time <- noun_onset + (i - 1) * pbin_dur
      end_time <- noun_onset + i * pbin_dur
      
      print(trial_id)
      print(participant)
      print(pbin_dur)
      print(paste("i:", i))
      print(paste("noun_onset:", noun_onset))
      print(paste("ending_offset:", ending_offset))
      print(paste("pbin_dur:", pbin_dur))
      print(paste("start_time:", noun_onset + 200 + (i - 1) * pbin_dur))
      print(paste("end_time:", noun_onset + 200 + i * pbin_dur))
      print(str(df_trial$TimeMS_adjusted))
      
      bins_in_interval <- df_trial %>%
        filter(TimeMS_adjusted >= start_time,
               TimeMS_adjusted < end_time) %>%
        pull(Bin) %>%
        sort()
      print(paste("bins in interval:", bins_in_interval))
      
      competitor_fix_count <- 0
      distractor_fix_count <- 0
      target_fix_count <- 0
      
      for (bin in bins_in_interval) {
        df_bin <- df_trial[df_trial$Bin == bin, ]
        
        if (df_bin$Competitor == 1) {
          competitor_fix_count = competitor_fix_count + 1
        }
        if (df_bin$Distractor == 1) {
          distractor_fix_count = distractor_fix_count + 1
        }
        if (df_bin$Target == 1) {
          target_fix_count = target_fix_count + 1
        }
      }
      
      competitor_fixP <- competitor_fix_count / pbin_dur
      distractor_fixP <- distractor_fix_count / pbin_dur
      target_fixP     <- target_fix_count / pbin_dur
      
      fix.rebinned <- rbind(
        fix.rebinned,
        data.frame(
          Trial = trial_id,
          Participant = participant,
          Pseudo_bin = i,
          Competitor_FixP = competitor_fixP,
          Distractor_FixP = distractor_fixP,
          Target_FixP = target_fixP,
          competitor_fix_count = competitor_fix_count,
          distractor_fix_count = distractor_fix_count,
          target_fix_count = target_fix_count,
          Pseudo_bin_duration = pbin_dur,
          Number_of_bins_in_interval = length(bins_in_interval)
        )
      )
    }
  }
}
```

# NOTE Should we take the time points to be the left edge, middle or right edge of the step function? Like how you'd take left-handed vs right-handed vs middle Riemann sums (right now it seems to be left edge)

# Another note: we are ignoring the code from this point until the new normalization part {style="color:red"}

##### Remove rows where the time bins are not 10ms {style="color:red"}

```{r eval=F, message=F}
fix.dat <- fix.dat %>% filter(Bin_duration == 10)
# TODO not sure if we should still do that
```

##### Add AreaFixated column {style="color:red"}

```{r eval=T, message=F}
fix.dat.full <- fix.dat.full %>%
  mutate(AreaFixated = case_when(
    Competitor == 1 ~ "Competitor",
    Distractor == 1 ~ "Distractor",
    Target == 1 ~ "Target",
    TRUE ~ "None"
  ))
# TODO no more AreaFixated because there can be multiple in one pseudo-bin. Need to find out if we can still pivot the data to a long format
# This is done later anyway, with fix.long
# for now we WILL do this step so the twovar step works
```

##### Reshape binomially encoded fixation data to a long format for plotting and analysis {style="color:green"}

In the original experiment, the eyetracker takes 25 samples in each 50ms bin, and records the proportion of samples that fall in each AOI. Our eyetracker takes 1 sample in each 10ms bin, and records *which* AOI the sample falls in (we already removed the trackloss bins, so the sample must be in one of the 3 AOIs, and binomially encoded as 0 or 1). *Update* actually, now that we added the missing rows back in, there could be rows where FixP=0 in all 3 columns And the pivoted data should be fine after normalization - we'll just have rows with more than 1 non-zero AOI columns

```{r eval=T, echo=T, message=F}
fix.long <- fix.dat.full %>%
  pivot_longer(
    cols = c(Target, Competitor, Distractor),
    names_to = "AOI",
    values_to = "FixP"
  )
```

##### Add GenderMatch column {style="color:red"}

```{r eval=F, echo=T, message=F}
# NOTE we didn't coalesce N_Gender_D, etc yet when we were filling in the missing rows
fix.long <- fix.long %>%
  mutate(GenderMatch = case_when(
    N_Gender_D == N_Gender_T ~ "D=T",
    N_Gender_D == N_Gender_C ~ "D=C"
  ))
```

##### Change `Participant`, `TOI`, `AreaFixated`, `SpeakerType`, `AOI`, `Phonol_Overlap_T` and `GenderMatch` to factor

```{r eval=F, echo=T, message=F}
fix.long = fix.long %>% mutate_at(vars(Participant, TOI, AreaFixated, SpeakerType, AOI, Phonol_Overlap_T, GenderMatch), as.factor)
```

##### Merge the non minimal pair values

```{r eval=F, echo=T, message=F}
fix.long$Phonol_Overlap <- fix.long$Phonol_Overlap_T %>% fct_collapse(NonMinPair = c("First Syll-unstressed", "First Syll-stressed"))
```

##### Change `Participant`, `TOI`, `AreaFixated`, `SpeakerType`, `AOI`, `Phonol_Overlap_T` and `GenderMatch` to factor

```{r eval=F, echo=T, message=F}
fix.long = fix.long %>% mutate_at(vars(Participant, TOI, SpeakerType, AOI, Determiner, Trackloss), as.factor)
```

##### Inspect the data after factoring

```{r eval=F, echo=T, message=F}
head(fix.long)
summary(fix.long)
```

##### Compute the mean, SD, SE and CI for `FixP` for each time bin, AOI, SpeakerType, Determiner and TOI

```{r eval=F, echo=T, message=F}
fix.means = Rmisc::summarySE(fix.long, measurevar='FixP', groupvars=c('TimeMS_adjusted','SpeakerType', 'AOI', 'Determiner'), na.rm=T)

head(fix.means)
summary(fix.means)
```

##### Compute Mean, etc. that takes Phonol_Overlap into account

```{r eval=F, echo=T, message=F}
fix.means.phonol = Rmisc::summarySE(fix.long, measurevar='FixP', groupvars=c('TimeMS_adjusted','SpeakerType', 'AOI', 'Determiner', 'Phonol_Overlap'), na.rm=T)

head(fix.means.phonol)
summary(fix.means.phonol)
```

##### Compute Mean, etc. that takes Phonol_Overlap AND GenderMatch into account

```{r eval=F, echo=T, message=F}
fix.means.fourgroups = Rmisc::summarySE(fix.long, measurevar='FixP', groupvars=c('TimeMS_adjusted','SpeakerType', 'AOI', 'Determiner', 'Phonol_Overlap', 'GenderMatch'), na.rm=T)

head(fix.means.fourgroups)
summary(fix.means.fourgroups)
```

##### Compute Mean, etc. that takes GenderMatch into account

```{r eval=F, echo=T, message=F}
fix.means.gendermatch = Rmisc::summarySE(fix.long, measurevar='FixP', groupvars=c('TimeMS_adjusted','SpeakerType', 'AOI', 'Determiner', 'GenderMatch'), na.rm=T)

head(fix.means.gendermatch)
summary(fix.means.gendermatch)
```

### 2. Plotting

##### Plot mean fixation proportion to each AOI with standard error, dividing by SpeakerType (ASI and HSS) and Determiner type (Definite and Possessive)

```{r eval=F, echo=T, message=F, fig.width=12}
fix.plot = ggplot() + 
  facet_wrap(~Determiner + SpeakerType, ncol=2,
             labeller = labeller(
              Determiner = c(
                "Definite" = "Definite determiner",
                "Possessive" = "Possessive determiner"
              ),
              SpeakerType = c(
                "1" = "ASI",
                "2" = "HSS"
              )
            )
             ) + theme_bw() + 
  geom_vline(xintercept = 0, linetype = "dashed", color = "black", linewidth = 0.8) +
  geom_line(data=fix.means,aes(x=TimeMS_adjusted, y=FixP, group=AOI, colour=AOI, lty=AOI), lwd=1.5) + # TODO use SE or CI here?
  geom_ribbon(data=fix.means,aes(x=TimeMS_adjusted,ymin=FixP-se,ymax=FixP+se,color=AOI,fill=AOI), linewidth=.2, alpha=.3, lty="dashed", show.legend=F)  +
  labs(y="Fixation Proportion", x="Time relative to the target word onset (ms)") + 
  
  scale_x_continuous(limits=c(-300, 1400),expand=c(0,0),breaks=seq(-300, 1400, 200)) +
  scale_color_manual('AOI', values=c("#af2013","#002e65","#ffac00")) +
  scale_fill_manual('AOI', values=c("#af2013","#002e65","#ffac00")) +
  scale_linetype_manual('AOI', values=c("solid","dotted", "dotdash")) +
  
  theme(
    text=element_text(size=20),
    legend.key.height=unit(.3,"in"),
    legend.key.width=unit(.6,"in"),
    panel.spacing=unit(1, "in")) 

fix.plot
```

##### Save the plot

```{r eval=F, message=F}
ggsave("plots/fixation_plot.png", width = 20, height = 15, dpi = 300)
```

##### Plot mean fixation proportion to each AOI with standard error, dividing by SpeakerType (ASI and HSS), Determiner type (Definite and Possessive) and Phonological Overlap (Minimal Pair and Non Minimal Pair)

```{r eval=F, echo=T, message=F, fig.width=20, fig.height=12}
fix.plot.phonol = ggplot() + 
  facet_wrap(~Determiner + SpeakerType + Phonol_Overlap, ncol=4,
             labeller = labeller(
              Determiner = c(
                "Definite" = "Definite determiner",
                "Possessive" = "Possessive determiner"
              ),
              SpeakerType = c(
                "1" = "ASI",
                "2" = "HSS"
              ),
              Phonol_Overlap = c(
                "Minimal Pair" = "Minimal Pair",
                "NonMinPair" = "Non Minimal Pair"
              )
            )
             ) + theme_bw() + 
  geom_vline(xintercept = 0, linetype = "dashed", color = "black", linewidth = 0.8) +
  geom_line(data=fix.means.phonol,aes(x=TimeMS_adjusted, y=FixP, group=AOI, colour=AOI, lty=AOI), lwd=1.5) + # TODO use SE or CI here?
  geom_ribbon(data=fix.means.phonol,aes(x=TimeMS_adjusted,ymin=FixP-se,ymax=FixP+se,color=AOI,fill=AOI), linewidth=.2, alpha=.3, lty="dashed", show.legend=F)  +
  labs(y="Fixation Proportion", x="Time relative to the target word onset (ms)") + 
  
  scale_x_continuous(limits=c(-300, 1400),expand=c(0,0),breaks=seq(-300, 1400, 200)) +
  scale_color_manual('AOI', values=c("#af2013","#002e65","#ffac00")) +
  scale_fill_manual('AOI', values=c("#af2013","#002e65","#ffac00")) +
  scale_linetype_manual('AOI', values=c("solid","dotted", "dotdash")) +
  
  theme(
    text=element_text(size=20),
    legend.key.height=unit(.3,"in"),
    legend.key.width=unit(.6,"in"),
    panel.spacing=unit(1, "in")) 

fix.plot.phonol
```

##### Save the plot

```{r eval=F, message=F}
ggsave("plots/fixation_plot_phonol.png", width = 40, height = 15, dpi = 300)
```

##### Plot mean fixation proportion to each AOI with standard error, dividing by SpeakerType (ASI and HSS), Determiner type (Definite and Possessive), Phonological Overlap (Minimal Pair and Non Minimal Pair) and GenderMatch (D=T and D=C)

```{r eval=F, echo=T, message=F, fig.width=20, fig.height=24}
fix.plot.fourgroups = ggplot() + 
  facet_wrap(~Determiner + SpeakerType + Phonol_Overlap + GenderMatch, ncol=4,
             labeller = labeller(
              Determiner = c(
                "Definite" = "Definite determiner",
                "Possessive" = "Possessive determiner"
              ),
              SpeakerType = c(
                "1" = "ASI",
                "2" = "HSS"
              ),
              Phonol_Overlap = c(
                "Minimal Pair" = "Minimal Pair",
                "NonMinPair" = "Non Minimal Pair"
              ),
              GenderMatch = c(
                "D=T" = "Distractor = Target",
                "D=C" = "Distractor = Competitor"
              )
            )
             ) + theme_bw() + 
  geom_vline(xintercept = 0, linetype = "dashed", color = "black", linewidth = 0.8) +
  geom_line(data=fix.means.fourgroups,aes(x=TimeMS_adjusted, y=FixP, group=AOI, colour=AOI, lty=AOI), lwd=1.5) + # TODO use SE or CI here?
  geom_ribbon(data=fix.means.fourgroups,aes(x=TimeMS_adjusted,ymin=FixP-se,ymax=FixP+se,color=AOI,fill=AOI), linewidth=.2, alpha=.3, lty="dashed", show.legend=F)  +
  labs(y="Fixation Proportion", x="Time relative to the target word onset (ms)") + 
  
  scale_x_continuous(limits=c(-300, 1400),expand=c(0,0),breaks=seq(-300, 1400, 200)) +
  scale_color_manual('AOI', values=c("#af2013","#002e65","#ffac00")) +
  scale_fill_manual('AOI', values=c("#af2013","#002e65","#ffac00")) +
  scale_linetype_manual('AOI', values=c("solid","dotted", "dotdash")) +
  
  theme(
    text=element_text(size=20),
    legend.key.height=unit(.3,"in"),
    legend.key.width=unit(.6,"in"),
    panel.spacing=unit(1, "in")) 

fix.plot.fourgroups
```

##### Save the plot

```{r eval=F, message=F}
ggsave("plots/fixation_plot_fourgroups.png", width = 40, height = 30, dpi = 300)
```

##### Plot mean fixation proportion to each AOI with standard error, dividing by SpeakerType (ASI and HSS), Determiner type (Definite and Possessive) and GenderMatch (D=T and D=C)

```{r eval=F, echo=T, message=F, fig.width=20, fig.height=12}
fix.plot.gendermatch = ggplot() + 
  facet_wrap(~Determiner + SpeakerType + GenderMatch, ncol=4,
             labeller = labeller(
              Determiner = c(
                "Definite" = "Definite determiner",
                "Possessive" = "Possessive determiner"
              ),
              SpeakerType = c(
                "1" = "ASI",
                "2" = "HSS"
              ),
              GenderMatch = c(
                "D=T" = "Distractor = Target",
                "D=C" = "Distractor = Competitor"
              )
            )
             ) + theme_bw() + 
  geom_vline(xintercept = 0, linetype = "dashed", color = "black", linewidth = 0.8) +
  geom_line(data=fix.means.gendermatch,aes(x=TimeMS_adjusted, y=FixP, group=AOI, colour=AOI, lty=AOI), lwd=1.5) + # TODO use SE or CI here?
  geom_ribbon(data=fix.means.gendermatch,aes(x=TimeMS_adjusted,ymin=FixP-se,ymax=FixP+se,color=AOI,fill=AOI), linewidth=.2, alpha=.3, lty="dashed", show.legend=F)  +
  labs(y="Fixation Proportion", x="Time relative to the target word onset (ms)") + 
  
  scale_x_continuous(limits=c(-300, 1400),expand=c(0,0),breaks=seq(-300, 1400, 200)) +
  scale_color_manual('AOI', values=c("#af2013","#002e65","#ffac00")) +
  scale_fill_manual('AOI', values=c("#af2013","#002e65","#ffac00")) +
  scale_linetype_manual('AOI', values=c("solid","dotted", "dotdash")) +
  
  theme(
    text=element_text(size=20),
    legend.key.height=unit(.3,"in"),
    legend.key.width=unit(.6,"in"),
    panel.spacing=unit(1, "in")) 

fix.plot.gendermatch
```

##### Save the plot

```{r eval=F, message=F}
ggsave("plots/fixation_plot_gendermatch.png", width = 40, height = 15, dpi = 300)
```

# Normalization take 2

### Combine Competitor and Distractor to create NonTarget columns and reduce the data set to two variables - this we do need to do

```{r eval=T, message=F}
fix.twovar <- fix.dat.full %>%
  mutate(NonTarget = Competitor + Distractor) %>%
  select(-Competitor, -Distractor)

fix.twovar <- fix.twovar %>%
  mutate(AreaFixated = ifelse(AreaFixated == "Target", "Target", "NonTarget"))

fix.twovar$AreaFixated <- as.factor(fix.twovar$AreaFixated)
# TODO right now we are NOT distinguishing between NonTarget and Trackloss
```

##### Pivot binomially encoded columns

```{r eval=T, message=F}
fix.twovar <- fix.twovar %>%
  pivot_longer(
    cols = c(Target, NonTarget),
    names_to = "AOI",
    values_to = "FixP"
  )

fix.twovar = fix.twovar %>% mutate_at(vars(AOI), as.factor)
```

### MIGHT BE DISCARDED create a step function for the average FixP of all participants for each trial. Right now, we are stretching the portion between determiner onset (time=0) and noun offset (ending offset). We *will* need to do this twice eventually, splitting the time-course data at *noun onset*. {style="color:red"}

##### Define windows for each trial

```{r eval=T, message=F}
# Assuming fix.twovar has: Participant, TOI, TimeMS_adjusted, FixP, AOI

# The number of pseudobins
N_PBINS <- 100

# Compute trial-specific windows using EndingOffset_adjusted
trial_windows <- fix.twovar %>%
  group_by(TOI) %>%
  summarise(
    trial_duration = max(EndingOffset_adjusted, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    window_list = map(trial_duration, function(dur) {
      edges <- seq(0, dur, length.out = N_PBINS + 1)
      tibble(start = head(edges, -1), end = tail(edges, -1))
    })
  ) %>%
  unnest(window_list)

# Group the fixation data by participant and trial)
# Should probably start with fix.twovar instead?
# fix.twovar.mean <- fix.twovar %>%
#   group_by(Participant, AOI, SpeakerType, Determiner, TOI, TimeMS_adjusted) %>%
#   summarise(FixP = mean(FixP, na.rm = TRUE), .groups = "drop")
# 
# summary(fix.twovar.mean)
```

##### Functions to calculate the AUC

```{r eval=T, message=F}
# Evaluate the AUC through integration; takes in a list of time points and corresponding FixPs
calc_auc_fancy <- function(time, fixp) {
  valid <- !is.na(time) & !is.na(fixp)
  time <- time[valid]
  fixp <- fixp[valid]
  if (length(time) < 2) return(NA_real_)
  f <- approxfun(x = time, y = fixp, method = "constant", rule = 2)
  # TODO rule=2 has something to do with constant extrapolation - extending the boundary values beyond the data range. Might be important in figuring out how to account for trackloss rows
  integrate(f, lower = min(time), upper = max(time))$value
  # TODO consider calculating the AUC manually as a weighted sum: If widths vary, AUC = sum(FixP[i] * (time[i+1]-time[i])).
}

# Calculate the AUC for a specified time interval within a dataframe; the dataframe must have columns of TimeMS_adjusted and FixP
calc_auc_interval <- function(df, start, end) {
  df_sub <- df %>%
    filter(TimeMS_adjusted >= start, TimeMS_adjusted < end)

  if (nrow(df_sub) < 2) { # if there's only one sample
    print(
      paste0(
        "Empty subset for window ",
        start, "-", end,
        " (n = ", nrow(df_sub), ")"
      )
    )
    return(NA_real_)
  }

  calc_auc_fancy(df_sub$TimeMS_adjusted, df_sub$FixP)
}
```

##### Test the AUC functions

```{r eval=T, message=F}
# Create a subset of the timecourse data
test_df <- fix.twovar %>%
  filter(Participant == "TA105",
         AOI == "Target",
         TOI == "Trial01-def-campesina_0")

# Check its time range
range(test_df$TimeMS_adjusted)

# test_df %>%
#   filter(TimeMS_adjusted >= 1000, TimeMS_adjusted < 1010)

calc_auc_interval(test_df, 1000, 1030)
```

##### Create test subsets

```{r eval=T, message=F}
# Create a test subset of fix.twovar with only one participant (and one of the participant's trials)
fix.twovar.participant = fix.twovar %>% filter(Participant == "TA105", TOI == "Trial01-def-campesina_0")

# Create a test subset of fix.twovar with only one trial (and all the trial's participants)
fix.twovar.trial = fix.twovar %>% filter(TOI == "Trial01-def-campesina_0")
```

##### Compute trial-specific AUCs for each participant × AOI

```{r eval=T, message=F}
auc_trial_windows <- pmap_dfr( # Uses purrr::pmap_dfr to iterate row-by-row over the tibble trial_windows, feeding each row’s columns as arguments to the anonymous function. It row-binds (_dfr) all returned tibbles into one data frame.
  trial_windows, # Must be a tibble whose columns match the function’s argument names exactly: here, TOI, trial_duration, start, end. Each row represents a single trial × window interval.
  function(TOI, trial_duration, start, end) { # The per-row function—receives scalars for the current trial id, its duration, and the window bounds, which are columns of trial_windows
    if (!TOI %in% fix.twovar.participant$TOI) return(tibble()) # Guard: if this trial id isn’t present in fix.twovar, skip this row by returning an empty tibble (which pmap_dfr will ignore).
    df_trial <- fix.twovar.participant %>% 
      mutate(trial_end = max(EndingOffset_adjusted, na.rm = TRUE)) %>%
      filter(TimeMS_adjusted >= 0, TimeMS_adjusted <= trial_end) # The !! (bang-bang) “injects” the scalar TOI value into the filter so it compares the column TOI to the current trial id. df_trial is a dataframe with all the rows from a trial in the original dataset
    if (nrow(df_trial) == 0) return(tibble()) # Another guard: if this trial has no rows (e.g., filtered away earlier), skip.

    df_trial %>% # For the current trial’s rows...
      group_by(Participant, AOI) %>%
      summarise(AUC = calc_auc_interval(pick(TimeMS_adjusted, FixP), start, end), .groups = "drop") %>%
      mutate(
        TOI = TOI,
        WindowStart = start,
        WindowEnd = end,
        WindowDur = end - start,
        AUC = AUC,
        AUC_prop = AUC / (end - start)
      )
  }
)

head(auc_trial_windows)
```

# I think the part below is the part that actually works

## 

##### Step 1: Create one step function per Participant × AOI × TOI

```{r eval=T, message=F}
step_functions <- fix.twovar %>%
  group_by(Participant, AOI, TOI) %>%
  summarise(
    f = list(approxfun(
      x = TimeMS_adjusted,
      y = FixP,
      method = "constant",
      rule = 2
    )),
    .groups = "drop"
  )
```

##### Example for one participant-trial (don't run this)

```{r eval=F, message=F}
# Example: AUC from 0–400 ms for one participant-trial
step_functions = step_functions %>% filter(Participant == "TA106", TOI == "Trial01-def-campesina_0") 
  # %>% mutate(AUC_0_5 = map_dbl(f, ~ integrate(.x, lower = 310, upper = 315)$value))
```

## Let's do it for determiners first

##### Step 2: Define windows for each trial

```{r eval=T, message=F}
# The number of pseudobins
N_PBINS <- 100

# Compute trial-specific windows using NounOnset_adjusted
trial_windows_det <- fix.twovar %>%
  group_by(TOI) %>%
  summarise(
    trial_duration = max(NounOnset_adjusted, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    window_list = map(trial_duration, function(dur) {
      edges <- seq(0, dur, length.out = N_PBINS + 1)
      tibble(start = head(edges, -1), end = tail(edges, -1))
    })
  ) %>%
  unnest(window_list)

# Compute AUCs for all windows per step function
# auc_by_window <- step_functions %>%
#   crossing(trial_windows) %>%
#   mutate(
#     AUC = map2_dbl(f, start, ~ integrate(.x, lower = .y, upper = .y + 400)$value),
#     WindowDur = 400,
#     AUC_prop = AUC / WindowDur
#   )

```

##### Step 3: Compute AUCs for all step functions

```{r}
auc_trial_windows_det <- step_functions %>%
  left_join(trial_windows_det, by = "TOI", relationship = "many-to-many") %>%
  mutate(
    AUC = pmap_dbl(
      list(f, start, end),
      function(fun, start, end) {
        integrate(fun, lower = start, upper = end)$value
      }
    ),
    WindowDur = end - start,
    AUC_prop = AUC / WindowDur
  )
```

##### Step 4: add pseudo-bin numbers

```{r eval=T, message=F}
auc_trial_windows_det <- auc_trial_windows_det %>%
  group_by(Participant, TOI) %>%
  arrange(start, AOI, .by_group = TRUE) %>%
  mutate(PBin = rep(1:100, each = length(unique(AOI)), length.out = n())) %>%
  ungroup()
```

##### Step 5: Compute average AUCs across participants (for smooth plotting)

```{r eval=T, message=F}
auc_summary_det <- auc_trial_windows_det %>%
  group_by(AOI, PBin) %>%
  summarise(
    mean_AUC = mean(AUC, na.rm = TRUE),
    se_AUC = sd(AUC, na.rm = TRUE) / sqrt(n()),
    mean_AUC_prop = mean(AUC_prop, na.rm = TRUE),
    se_AUC_prop = sd(AUC_prop, na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  )
```

##### Step 6: Plot the summary of AUC proportions across trials and participants

```{r eval=T, message=F}
ggplot(auc_summary_det, aes(x = PBin, y = mean_AUC_prop, color = AOI, group = AOI)) +
  geom_line(size = 1.2) +
  geom_ribbon(aes(ymin = mean_AUC_prop - se_AUC_prop, ymax = mean_AUC_prop + se_AUC_prop, fill = AOI),
              alpha = 0.2, color = NA) +
  scale_color_manual(values = c("Target" = "#ffac00", "NonTarget" = "#002e65")) +
  scale_fill_manual(values = c("Target" = "#ffac00", "NonTarget" = "#002e65")) +
  labs(
    title = "Mean AUC over Time for Determiners",
    x = "Pseudo-bin",
    y = "AUC proportions",
    color = "AOI",
    fill = "AOI"
  ) +
  theme_minimal(base_size = 16)
```

## Now do it for noun portions

##### Step 2: Define windows for each trial

```{r eval=T, message=F}
# The number of pseudobins
N_PBINS <- 100

trial_windows_noun <- fix.twovar %>%
  group_by(TOI) %>%
  summarise(
    start_anchor = min(NounOnset_adjusted, na.rm = TRUE),
    end_anchor   = max(EndingOffset_adjusted, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    window_list = map2(start_anchor, end_anchor, function(start, end) {
      edges <- seq(start, end, length.out = N_PBINS + 1)
      tibble(start = head(edges, -1), end = tail(edges, -1))
    })
  ) %>%
  unnest(window_list)

```

##### Step 3: Compute AUCs for all step functions

```{r}
auc_trial_windows_noun <- step_functions %>%
  left_join(trial_windows_noun, by = "TOI", relationship = "many-to-many") %>%
  mutate(
    AUC = pmap_dbl(
      list(f, start, end),
      function(fun, start, end) {
        integrate(fun, lower = start, upper = end)$value
      }
    ),
    WindowDur = end - start,
    AUC_prop = AUC / WindowDur
  )
```

##### Step 4: add pseudo-bin numbers

```{r eval=T, message=F}
auc_trial_windows_noun <- auc_trial_windows_noun %>%
  group_by(Participant, TOI) %>%
  arrange(start, AOI, .by_group = TRUE) %>%
  mutate(PBin = rep(1:100, each = length(unique(AOI)), length.out = n())) %>%
  ungroup()
```

##### Step 5: Compute average AUCs across participants (for smooth plotting)

```{r eval=T, message=F}
auc_summary_noun <- auc_trial_windows_noun %>%
  group_by(AOI, PBin) %>%
  summarise(
    mean_AUC = mean(AUC, na.rm = TRUE),
    se_AUC = sd(AUC, na.rm = TRUE) / sqrt(n()),
    mean_AUC_prop = mean(AUC_prop, na.rm = TRUE),
    se_AUC_prop = sd(AUC_prop, na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  )
```

##### Step 6: Plot the summary of AUC proportions across trials and participants

```{r eval=T, message=F}
ggplot(auc_summary_noun, aes(x = PBin, y = mean_AUC_prop, color = AOI, group = AOI)) +
  geom_line(size = 1.2) +
  geom_ribbon(aes(ymin = mean_AUC_prop - se_AUC_prop, ymax = mean_AUC_prop + se_AUC_prop, fill = AOI),
              alpha = 0.2, color = NA) +
  scale_color_manual(values = c("Target" = "#ffac00", "NonTarget" = "#002e65")) +
  scale_fill_manual(values = c("Target" = "#ffac00", "NonTarget" = "#002e65")) +
  labs(
    title = "Mean AUC over Time for Nouns",
    x = "Pseudo-bin",
    y = "AUC proportions",
    color = "AOI",
    fill = "AOI"
  ) +
  theme_minimal(base_size = 16)
```

##### Step 7: Concatenate the two plots

```{r}
# Step 1: Shift the noun pseudo-bins
auc_summary_noun_shifted <- auc_summary_noun %>%
  mutate(PBin = PBin + 100)
```

```{r}
# Step 2: Combine determiner + noun data
auc_combined <- bind_rows(
  auc_summary_det,
  auc_summary_noun_shifted
)
```

```{r}
# Step 3: Plot (almost exactly your original code)
ggplot(auc_combined,
       aes(x = PBin, y = mean_AUC_prop, color = AOI, group = AOI)) +
  geom_line(linewidth = 1.2) +
  geom_ribbon(
    aes(ymin = mean_AUC_prop - se_AUC_prop,
        ymax = mean_AUC_prop + se_AUC_prop,
        fill = AOI),
    alpha = 0.2,
    color = NA
  ) +
  scale_color_manual(values = c("Target" = "#ffac00",
                                "NonTarget" = "#002e65")) +
  scale_fill_manual(values = c("Target" = "#ffac00",
                               "NonTarget" = "#002e65")) +
  labs(
    title = "Mean AUC over Time: Determiner → Noun",
    x = "Pseudo-bin (normalized time)",
    y = "AUC proportion",
    color = "AOI",
    fill = "AOI"
  ) +
  theme_minimal(base_size = 16)

```

##### Plots of step functions for individual participant-trials

```{r eval=T, message=F}
f_example <- step_functions %>%
  filter(Participant == "TA106", AOI == "Target", TOI == "Trial01-def-campesina_0") %>%
  pull(f) %>%
  .[[1]]

t_seq <- seq(-300, 1400, by = 10)

plot(t_seq, f_example(t_seq),
     type = "s", col = "#ffac00", lwd = 2,
     main = "Step Function for TA106 – Trial01-def-campesina_0",
     xlab = "Time (ms, adjusted)", ylab = "Fixation Probability")
```

```{r eval=T, message=F}
# auc_by_participant_trial <- fix.twovar %>%
#   group_by(Participant, TOI, AOI, SpeakerType, Determiner) %>% # here, we group by the variables that we will compare between later
#   summarise(AUC = sum(FixP) * 10)
# 
# head(auc_by_participant_trial)
# 
# 
# intervals <- data.frame(
#   start = c(0, 400, 800),
#   end   = c(400, 800, 1200),
#   label = c("0–400", "400–800", "800–1200")
# )

# fix.intervalled <- fix.twovar %>%
#   mutate(
#     TimeWindow = case_when(
#       TimeMS_adjusted >= 0   & TimeMS_adjusted < 400  ~ "0–400",
#       TimeMS_adjusted >= 400 & TimeMS_adjusted < 800  ~ "400–800",
#       TimeMS_adjusted >= 800 & TimeMS_adjusted <= 1200 ~ "800–1200",
#       TRUE ~ NA_character_
#     )
#   ) %>%
#   filter(!is.na(TimeWindow))

# auc_by_window <- fix.intervalled %>%
#   group_by(Participant, AOI, TOI, TimeWindow) %>%
#   summarise(
#     AUC = calc_auc_fancy(TimeMS_adjusted, FixP),
#     .groups = "drop"
#   )



# Compute trial-specific AUC windows
# auc_trial_windows <- trial_windows %>%
#   select(TOI, window_list) %>%
#   unnest(window_list) %>%
#   group_by(TOI, start, end) %>%
#   group_modify(~{
#     df_trial <- fix.twovar %>% filter(TOI == .y$TOI)
#     
#     df_trial %>%
#       group_by(Participant, AOI) %>%
#       summarise(
#         AUC = calc_auc_interval(cur_data(), .x$start, .x$end),
#         .groups = "drop"
#       ) %>%
#       mutate(WindowStart = .x$start, WindowEnd = .x$end)
#   }) %>%
#   ungroup()
```

```{r eval=T, message=F}
# Define a function to make a step function for each participant and trial
# make_stepfun <- function(time, fixp) {
#   approxfun(x = time, y = fixp, method = "constant", rule = 2)
# }
# 
# # Define a function to calculate the area under the step function for any time interval 
# calc_auc <- function(time, fixp, start, end) {
#   f <- make_stepfun(time, fixp)
#   integrate(f, lower = start, upper = end)$value
# }


unique(fix.twovar$Participant)[1:10]
unique(fix.twovar$TOI)[1:10]
levels(fix.twovar$AOI)

# Example subset
subj_data <- fix.twovar %>%
  filter(Participant == "TA108", AOI == "Target",
         TOI == "Trial09-pos-cocinera_0") %>%
  arrange(TimeMS_adjusted)

head(subj_data)

# calc_auc_fancy(subj_data$TimeMS_adjusted, subj_data$FixP)
# calc_auc_interval(subj_data, 200, 800)

# -> columns: TimeMS_adjusted, FixP (0/1)
# 
f_step <- approxfun(subj_data$TimeMS_adjusted, subj_data$FixP,
                    method = "constant", rule = 2)

# Generate continuous time points for plotting
t_seq <- seq(min(subj_data$TimeMS_adjusted),
             max(subj_data$TimeMS_adjusted),
             by = 10)

plot(t_seq, f_step(t_seq),
     type = "s", lwd = 2, col = "#ffac00",
     xlab = "Time (ms)",
     ylab = "Fixation Probability",
     main = paste("Step Function", unique(subj_data$Participant)))




# # Apply the two functions across participants and trials
# auc_summary <- fix.twovar.mean %>%
#   group_by(Participant, AOI, SpeakerType, Determiner) %>%
#   summarise(AUC = calc_auc(TimeMS_adjusted, FixP),
#             .groups = "drop")

# summary(auc_summary)
```

##### Merge the non minimal pair values

```{r eval=F, echo=T, message=F}
fix.twovar$Phonol_Overlap <- fix.twovar$Phonol_Overlap_T %>% fct_collapse(NonMinPair = c("First Syll-unstressed", "First Syll-stressed"))

summary(fix.twovar)
head(fix.twovar)
```

##### Compute mean fixation proportions for the two-variable version

```{r eval=T, echo=T, message=F}
fix.means.twovar = Rmisc::summarySE(fix.twovar, measurevar='FixP', groupvars=c('TimeMS_adjusted','SpeakerType', 'AOI', 'Determiner'), na.rm=T)

fix.means.twovar = fix.means.twovar %>% mutate_at(vars(SpeakerType, Determiner), as.factor)

summary(fix.means.twovar)
head(fix.means.twovar)
```

##### Plot mean fixation proportion vs. time with the two-variable version of the data

```{r eval=T, echo=T, message=F, fig.width = 12}
fix.plot.twovar = ggplot() + 
  facet_wrap(~Determiner + SpeakerType, ncol=2,
             labeller = labeller(
              Determiner = c(
                "Definite" = "Definite determiner",
                "Possessive" = "Possessive determiner"
              ),
              SpeakerType = c(
                "1" = "ASI",
                "2" = "HSS"
              )
            )
             ) + theme_bw() + 
  geom_vline(xintercept = 0, linetype = "dashed", color = "black", linewidth = 0.8) +
  geom_line(data=fix.means.twovar,aes(x=TimeMS_adjusted, y=FixP, group=AOI, colour=AOI, lty=AOI), lwd=1.5) +
  geom_ribbon(data=fix.means.twovar,aes(x=TimeMS_adjusted,ymin=FixP-se,ymax=FixP+se,color=AOI,fill=AOI), linewidth=.2, alpha=.3, lty="dashed", show.legend=F)  +
  labs(y="Fixation Proportion", x="Time relative to the target word onset (ms)") + 
  
  scale_x_continuous(limits=c(-300, 1400),expand=c(0,0),breaks=seq(-300, 1400, 200)) +
  scale_color_manual('AOI', values=c("#969798","#ffac00")) +
  scale_fill_manual('AOI', values=c("#969798","#ffac00")) +
  scale_linetype_manual('AOI', values=c("solid","dotted")) +
  
  theme(
    text=element_text(size=20),
    legend.key.height=unit(.3,"in"),
    legend.key.width=unit(.6,"in"),
    panel.spacing=unit(1, "in")) 

fix.plot.twovar
```

##### Save the plot

```{r eval=T, message=F}
ggsave("plots/fixation_plot_twovar.png", width = 20, height = 15, dpi = 300)
```

### 3. DPA

Here we will run a by-participant analysis testing the divergence between the Target and NonTarget AOIs.

##### Select time window and create stratification variables

```{r eval=T, echo=T, message=F}
div.dat = fix.twovar %>%  
  filter(1400>=TimeMS_adjusted & TimeMS_adjusted>=-300) %>% 
  dplyr::mutate(StrataVars=paste(Participant, AOI, TimeMS_adjusted, sep='')) %>% 
  mutate_at(vars(SpeakerType, StrataVars),as.factor) %>% 
  droplevels()

summary(div.dat)
head(div.dat)
```

##### Create two subsets based on the type of determiner

```{r eval=T, echo=T, message=F}
div.dat.def <- div.dat %>% filter(Determiner == "Definite")
div.dat.pos <- div.dat %>% filter(Determiner == "Possessive")
```

##### Define bootstrap functions to compute a divergence point for the target vs. unrelated conditions

Number of consecutive statistically significant time bins required to establish divergence, and number of bootstrap iterations

```{r eval=T, echo=T, message=F}
Nbins = 20
Niter = 100L
```

Bootstrap function using lmer

```{r eval=T, echo=T, message=F}
boot_L1L2_lmer = function(original_data, resample_indices){
  dat = original_data[resample_indices, ]
  
  # a statistical test at each timepoint for each group
  test_g1 = dat %>% # t-test for L1 group
    subset(SpeakerType == 1) %>% group_by(TimeMS_adjusted) %>%
    dplyr::summarise(t = summary(lmer(FixP ~ AOI + (1|Participant)))$coefficients["AOITarget", "t value"])

  test_g2 = dat %>% # t-test for L2 group
    subset(SpeakerType == 2) %>% group_by(TimeMS_adjusted) %>%
    dplyr::summarise(t = summary(lmer(FixP ~ AOI + (1|Participant)))$coefficients["AOITarget", "t value"])

  # return a TRUE/FALSE vector of significant positive t-scores
  # (positive means more looks to the target than unrelated)
  t_g1 = test_g1$t > 1.96
  t_g2 = test_g2$t > 1.96
  
  # create empty vectors to store onsets
  onset_g1 = onset_g2 = c()
  
  # find the index of the earliest run of sequential TRUEs; threshold is Nbins
  for (i in 1:(length(t_g2)-Nbins)) {
    onset_g1[i] = sum(t_g1[i:(i+Nbins-1)]) == Nbins
    onset_g2[i] = sum(t_g2[i:(i+Nbins-1)]) == Nbins
  }
  
  # find the difference between onsets
  delta_g1g2 = which(onset_g2)[1] - which(onset_g1)[1]
  
  # print 
  # note: the bootstrap returns the indices of the respective timepoints, not absolute times. 
  # The annotations to the right of each index (e.g. t[,1]) indicate where in the boot object the bootstrapped onset distributions can be found.
  c(delta_g1g2,         # onset difference L1 vs. L2 t[,1]
    which(onset_g1)[1], # onset bin for looks to target L1 t[,2]
    which(onset_g2)[1])  # onset bin for looks to target L2 t[,3]
}

```

Run the lmer bootstrap on the two subsets by determiner type

```{r eval=F, echo=T, message=F}
bootres_L1L2_lmer_def = boot::boot(
  data = div.dat.def,  # data set to bootstrap       
  statistic = boot_L1L2_lmer,  # bootstrap function      
  strata = div.dat.def$StrataVars, # stratification variable 
  R = Niter)  # number of iterations

bootres_L1L2_lmer_pos = boot::boot(
  data = div.dat.pos,  # data set to bootstrap       
  statistic = boot_L1L2_lmer,  # bootstrap function      
  strata = div.dat.pos$StrataVars, # stratification variable 
  R = Niter)  # number of iterations

bootres_L1L2_lmer_def
bootres_L1L2_lmer_pos
```

Save the lmer results

```{r eval=F, echo=T, message=F}
save(bootres_L1L2_lmer_def, file="DPA_results.lmer.def.rds")
save(bootres_L1L2_lmer_pos, file="DPA_results.lmer.pos.rds")
```

##### Running the lmer bootstrap by trial

This ended up yielding mostly NAs due to the small sample size.

Define a function that runs the bootstrap function for all the trials

```{r eval=F, echo=T, message=F}
run_boot_per_trial <- function(toi_id, df, Niter = 100, strata_col = "StrataVars") {
  trial_data <- df %>% filter(TOI == toi_id) %>% droplevels()

  boot_result <- boot::boot(
    data = trial_data,
    statistic = boot_L1L2_glm,  # your divergence function
    strata = trial_data[[strata_col]],
    R = Niter
  )

  # Extract metadata from the first row (assuming Determiner etc. are constant per trial)
  trial_info <- trial_data %>%
    slice(1) %>%
    select(TOI, Determiner, Phonol_Overlap, GenderMatch)  # add more if needed

  list(boot = boot_result, info = trial_info)
}
```

Run the function that runs the bootstrap function

```{r eval=F, echo=T, message=F}
trial_ids <- unique(div.dat$TOI)

# Store both the bootstrap and metadata
boot_trials <- lapply(trial_ids, function(toi) run_boot_per_trial(toi, div.dat, Niter = 100))
names(boot_trials) <- trial_ids
```

Extract divergence values + metadata into a summary dataframe

```{r eval=F, echo=T, message=F}
# Function to extract means and metadata
extract_div_summary <- function(result) {
  bootres <- result$boot
  info <- result$info

  tibble::tibble(
    TOI = info$TOI,
    Determiner = info$Determiner,
    Phonol_Overlap = info$Phonol_Overlap,
    GenderMatch = info$GenderMatchTarget_D,
    ASI_divergence = mean(bootres$t[,2], na.rm = TRUE) * 10 - 10,
    HSS_divergence = mean(bootres$t[,3], na.rm = TRUE) * 10 - 10
  )
}

# Build summary
div_summary_df <- purrr::map_dfr(boot_trials, extract_div_summary)
```

Bootstrap function using glm (generalized linear model)

```{r eval=F, echo=T, message=F}
boot_L1L2_glm = function(original_data, resample_indices){
  dat = original_data[resample_indices, ]
  
  # a statistical test at each timepoint for each group
  test_g1 = dat %>% # t-test for L1 group
    subset(SpeakerType == 1) %>% group_by(TimeMS_adjusted) %>%
    dplyr::summarise(t = summary(glm(FixP ~ AOI, family = binomial))$coefficients["AOITarget", "z value"])

  test_g2 = dat %>% # t-test for L2 group
    subset(SpeakerType == 2) %>% group_by(TimeMS_adjusted) %>%
    dplyr::summarise(t = summary(glm(FixP ~ AOI, family = binomial))$coefficients["AOITarget", "z value"])

  # return a TRUE/FALSE vector of significant positive t-scores
  # (positive means more looks to the target than unrelated)
  t_g1 = test_g1$t > 1.96
  t_g2 = test_g2$t > 1.96
  
  # create empty vectors to store onsets
  onset_g1 = onset_g2 = c()
  
  # find the index of the earliest run of sequential TRUEs; threshold is Nbins
  for (i in 1:(length(t_g2)-Nbins)) {
    onset_g1[i] = sum(t_g1[i:(i+Nbins-1)]) == Nbins
    onset_g2[i] = sum(t_g2[i:(i+Nbins-1)]) == Nbins
  }
  
  # find the difference between onsets
  delta_g1g2 = which(onset_g2)[1] - which(onset_g1)[1]
  
  # print 
  # note: the bootstrap returns the indices of the respective timepoints, not absolute times. 
  # The annotations to the right of each index (e.g. t[,1]) indicate where in the boot object the bootstrapped onset distributions can be found.
  c(delta_g1g2,         # onset difference L1 vs. L2 t[,1]
    which(onset_g1)[1], # onset bin for looks to target L1 t[,2]
    which(onset_g2)[1])  # onset bin for looks to target L2 t[,3]
}

```

Run the glm bootstrap on the two subsets by determiner type

```{r eval=F, echo=T, message=F}
bootres_L1L2_glm_def = boot::boot(
  data = div.dat.def,  # data set to bootstrap       
  statistic = boot_L1L2_glm,  # bootstrap function      
  strata = div.dat.def$StrataVars, # stratification variable 
  R = Niter)  # number of iterations

bootres_L1L2_glm_pos = boot::boot(
  data = div.dat.pos,  # data set to bootstrap       
  statistic = boot_L1L2_glm,  # bootstrap function      
  strata = div.dat.pos$StrataVars, # stratification variable 
  R = Niter)  # number of iterations

bootres_L1L2_glm_def
bootres_L1L2_glm_pos
```

Save the glm results

```{r eval=F, echo=T, message=F}
save(bootres_L1L2_glm_def, file="DPA_results.glm.def.rds")
save(bootres_L1L2_glm_pos, file="DPA_results.glm.pos.rds")
```

##### Extracting and plotting the divergence points

Load the lmer results

```{r eval=T, echo=T, message=F}
load("DPA_results.lmer.def.rds")
load("DPA_results.lmer.pos.rds")
```

Load the glm results

```{r eval=F, echo=T, message=F}
load("DPA_results.glm.def.rds")
load("DPA_results.glm.pos.rds")
```

Define function to extract divergence points and confidence intervals from bootstrapping results

```{r eval=T, echo=T, message=F}
get_divergence_ms <- function(bootres, index) {
  # index = 2 for ASI, 3 for HSS
  mean_bin <- mean(bootres$t[, index], na.rm = TRUE)
  ci_bin <- boot::boot.ci(bootres, index = index, type = "perc")$percent[4:5]
  
  mean_ms <- (mean_bin - 1) * 10
  ci_ms <- (ci_bin - 1) * 10
  
  list(mean = mean_ms, ci = ci_ms)
}
```

Run the function on all the results

```{r eval=T, echo=T, message=F}
# Definite determiner
def_asi <- get_divergence_ms(bootres_L1L2_lmer_def, index = 2)
def_hss <- get_divergence_ms(bootres_L1L2_lmer_def, index = 3)

# Possessive determiner
pos_asi <- get_divergence_ms(bootres_L1L2_lmer_pos, index = 2)
pos_hss <- get_divergence_ms(bootres_L1L2_lmer_pos, index = 3)
```

Create dataframe for plotting

```{r eval=T, echo=T, message=F}
div_df <- tibble::tibble(
  Determiner = c("Definite", "Definite", "Possessive", "Possessive"),
  SpeakerType = c("1", "2", "1", "2"),
  TimeMS = c(def_asi$mean, def_hss$mean, pos_asi$mean, pos_hss$mean),
  CI_low = c(def_asi$ci[1], def_hss$ci[1], pos_asi$ci[1], pos_hss$ci[1]),
  CI_high = c(def_asi$ci[2], def_hss$ci[2], pos_asi$ci[2], pos_hss$ci[2])
)

div_df = div_df %>% mutate_at(vars(SpeakerType, Determiner), as.factor)

div_df <- div_df %>%
  mutate(label = paste0(TimeMS, " ms"))
```

Add divergence points and their confidence intervals to the two-variable plots

```{r eval=T, echo=T, message=F, fig.width = 14}
fix.plot.twovar +
  geom_point(data = div_df,
             aes(x = TimeMS, y = 0.48),
             shape = 18, size = 4, color = "black") +
  geom_errorbarh(data = div_df,
                 aes(xmin = CI_low, xmax = CI_high, y = 0.48),
                 height = 0.02, color = "black", linewidth = 0.6) +
  geom_text(data = div_df,
          aes(x = TimeMS - 50, y = 0.55, label = label), 
          size = 10, fontface="bold")
```

Save the plots

```{r eval=T, message=F}
ggsave("plots/fixation_plot_twovar_divpts.png", width = 20, height = 15, dpi = 300)
```

### 4. Group comparison

We can compute a p-value for the group comparison by creating a bootstrap distribution of the null hypothesis. To do this, we will first pool the original data from both groups, randomly assign group labels, and then estimate a difference in divergence point. We will repeat this many times to obtain a distribution of divergence point that could be expected under the null hypothesis. We will obtain the p-value by calculating the proportion of samples from this null distribution that are larger than the observed difference in divergence point in the empirical data.

Below, we will calculate a p-value for the group comparison (ASI vs. HSS group). We will first define a bootstrap function.

##### Define the bootstrap function for group comparison (only lmer so far)

```{r eval=T, echo=T, message=F}
boot_L1L2_pval = function(original_data, resample_indices){
  dat_resample = original_data[resample_indices, ] # resample the data
  
  dat = dat_resample %>% 
    group_by(Participant, AOI, TimeMS_adjusted) %>% 
    transform(SpeakerType=sample(SpeakerType,replace=F)) %>% # randomly assign group labels 
    ungroup() #%>% 
  
  # a statistical test at each timepoint for each group
  test_g1 = dat %>% # t-test for L1 group
    subset(SpeakerType == 1) %>% group_by(TimeMS_adjusted) %>%
    dplyr::summarise(t = summary(lmer(FixP ~ AOI + (1|Participant)))$coefficients["AOITarget", "t value"])

  test_g2 = dat %>% # t-test for L2 group
    subset(SpeakerType == 2) %>% group_by(TimeMS_adjusted) %>%
    dplyr::summarise(t = summary(lmer(FixP ~ AOI + (1|Participant)))$coefficients["AOITarget", "t value"])
  
  # return a TRUE/FALSE vector of significant positive t-scores  
  # (positive means more looks to the target than unrelated)
  t_g1 = test_g1$t > 1.96
  t_g2 = test_g2$t > 1.96
  
  # create empty vectors to store onsets
  onset_g1 = onset_g2 = c()
  
  # find the index of the earliest run of 4 sequential TRUEs 
  for (i in 1:(length(t_g2)-Nbins)) { 
    onset_g1[i] = sum(t_g1[i:(i+Nbins-1)]) == Nbins
    onset_g2[i] = sum(t_g2[i:(i+Nbins-1)]) == Nbins
  }
  
  # find the difference between onsets
  delta_g1g2 = which(onset_g2)[1] - which(onset_g1)[1]
  
  # print 
  delta_g1g2  # onset difference L1 vs. L2 t[,1]
}

```

##### Run the lmer bootstrap on the two subsets by determiner type

```{r eval=F, echo=T, message=F}
bootres_L1L2_pval_def = boot::boot(
  data = div.dat.def,  # data set to bootstrap       
  statistic = boot_L1L2_pval,  # bootstrap function      
  strata = div.dat.def$StrataVars, # stratification variable 
  R = Niter)  # number of iterations        

bootres_L1L2_pval_pos = boot::boot(
  data = div.dat.pos,  # data set to bootstrap       
  statistic = boot_L1L2_pval,  # bootstrap function      
  strata = div.dat.pos$StrataVars, # stratification variable 
  R = Niter)  # number of iterations   

bootres_L1L2_pval_def
bootres_L1L2_pval_pos
```

##### Save the group comparison results

```{r eval=F, echo=T, message=F}
save(bootres_L1L2_pval, file="DPA_results.pval.def.rds")
save(bootres_L1L2_pval, file="DPA_results.pval.pos.rds")
```

##### If we already have saved results, we can load them.

```{r eval=F, echo=T, message=F}
load("DPA_results.pval.def.rds")
load("DPA_results.pval.pos.rds")
```

##### Compute the p-values

```{r eval=F, echo=T, message=F}
pval.def <- round(mean(bootres_L1L2_pval_def$t[,1]>=bootres_L1L2_lmer_def$t0[1], na.rm=T), 5)        
pval.pos <- round(mean(bootres_L1L2_pval_pos$t[,1]>=bootres_L1L2_lmer_pos$t0[1], na.rm=T), 5) 

pval.def
pval.pos
```
