---
title: "Normalized_DPA"
output: html_document
date: "2026-05-30"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. Data preparation

```{r eval=T, echo=T, message=F}
rm(list = ls()) # Clear the workspace
options(scipen = 999) # Turn off scientific notation
```

## Packages

```{r eval=T, message=F}
require(Rmisc)
require(tidyverse)
require(ggplot2)
require(boot)
require(dplyr)
require(lme4)
require(purrr)
require(languageserver)
```

## Package versions

```{r}
installed.packages()[names(sessionInfo()$otherPkgs), "Version"]
```

## 1.1. Define constants

```{r}
# Window of interest (in milliseconds)
WINDOW_START <- -300
WINDOW_END <- 2000

# Number of pseudobins for normalization
N_PBINS = 100

# Number of consecutive statistically significant time bins required to establish divergence
Nbins = 20

# Number of bootstrap iterations
Niter = 100L
```

## 1.2. Read the time-course output data of the eye-tracker from file

```{r}
fix.dat = read_delim("data-2025-03-12/df_cleaned.csv", delim=',')
fix.dat = fix.dat %>% mutate_at(vars(Timeline), as.factor)
# all the trackloss bins have been removed - in any given row, there is a fixation in one of the 3 AreaFixated 

summary(fix.dat)
head(fix.dat)
```

## 1.3. Fill in the missing time points and add a trackloss column

```{r}
# Make a list of relevant trial variables
trial_vars <- c(
  "SpeakerType",
  "Determiner",
  "Phonol_Overlap_T",
  "NounOnset_adjusted",
  "EndingOffset_adjusted"
)

# Create rows for the missing time points
fix.dat.full <- fix.dat %>%
  group_by(Participant, TOI) %>%
  complete(
    TimeMS_adjusted = seq(
      min(TimeMS_adjusted),
      max(TimeMS_adjusted),
      by = 10
    )
  ) %>%
  
  # Fill in the values for the trial variables
  fill(all_of(trial_vars), .direction = "downup") %>%
  
  # Add binary fixation proportions of 0 to each AOI
  mutate(
    across(
      c(Target, Competitor, Distractor),
      ~ coalesce(.x, 0)
    )
  ) %>%
  ungroup()
```

## 1.4. Factorize relevant variables and drop extra columns

```{r}
fix.dat.full <- fix.dat.full %>%
  
  # Change independent variables to factors
  mutate(
    TOI = factor(TOI),
    Participant = factor(Participant),
    Determiner = factor(Determiner),
    SpeakerType = factor(SpeakerType),
    Phonol_Overlap_T = factor(Phonol_Overlap_T)
  ) %>%
  
  # Drop the extra columns
  select(
    Participant,
    TOI,
    TimeMS_adjusted,
    SpeakerType,
    Determiner,
    Phonol_Overlap_T,
    Target,
    Competitor,
    Distractor,
    NounOnset_adjusted,
    EndingOffset_adjusted
  ) %>%
  
  # Collapse the two types of non-minimal pairs in the phonological overlap variable
  mutate(
    Phonol_Overlap_T = forcats::fct_collapse(
      Phonol_Overlap_T,
      "Non Minimal Pair" = c(
        "First Syll-stressed",
        "First Syll-unstressed"
      )
    )
  ) %>%
  
  # Add an AreaFixated column
  mutate(
    AreaFixated = case_when(
      Target == 1 ~ "Target",
      Competitor == 1 ~ "Competitor",
      Distractor == 1 ~ "Distractor",
      TRUE ~ "Trackloss"
    )
  )

summary(fix.dat.full)
```

### Sanity check plot (step function plot for one participant and TOI)

```{r eval=F}
trial_id <- "Trial02-def-perro_0"
participant_id <- "TA105"

df_plot <- fix.dat.full %>%
  filter(Participant == participant_id, TOI == trial_id) %>%
  select(TimeMS_adjusted, Target, Competitor, Distractor) %>%
  pivot_longer(cols = c(Target, Competitor, Distractor),
               names_to = "AOI", values_to = "Fixation")

library(ggplot2)

vline_df <- tibble::tibble(
  time = c(108, 551),
  event = c("Noun onset", "Ending offset")
)

ggplot(df_plot, aes(x = TimeMS_adjusted, y = Fixation, color = AOI)) +
  geom_step(size = 0.8) +
  geom_vline(
    data = vline_df,
    aes(xintercept=time, linetype=event),
    color="black",
    linewidth=0.8
  ) +
  scale_x_continuous(limits = c(WINDOW_START, WINDOW_END)) +
  scale_y_continuous(limits = c(0, 1)) +
  labs(title = "Binary Fixations by AOI",
       subtitle = "Participant TA105 – Trial02-def-perro_0",
       x = "Time (ms)",
       y = "Fixated (0/1)") +
  theme_minimal()
```

### Average noun onset and ending offset

```{r eval=F}
fix.dat.full %>%
  summarise(avg_onset = mean(unique(NounOnset_adjusted), na.rm = TRUE))
fix.dat.full %>%
  summarise(avg_offset = mean(unique(EndingOffset_adjusted), na.rm = TRUE))
```

### Calculate the proportion of trackloss for each trial

(why are some trackloss proportions 1???)

```{r}
library(dplyr)

trackloss_summary <- fix.dat.full %>%
  
  # mark trackloss rows
  mutate(trackloss = AreaFixated == "Trackloss") %>%
  
  # classify time window
  mutate(window = case_when(
    TimeMS_adjusted >= 0 & TimeMS_adjusted < NounOnset_adjusted ~ "determiner_portion",
    TimeMS_adjusted >= NounOnset_adjusted & TimeMS_adjusted <= EndingOffset_adjusted ~ "noun_portion",
    TRUE ~ NA_character_
  )) %>%
  
  # keep only relevant time bins
  filter(!is.na(window)) %>%
  
  # compute proportion
  group_by(Participant, TOI, window) %>%
  summarise(
    prop_trackloss = mean(trackloss),
    n_samples = n(),
    .groups = "drop"
  )

avg_trackloss <- trackloss_summary %>%
  group_by(window) %>%
  summarise(
    avg_prop_trackloss = mean(prop_trackloss),
    .groups = "drop"
  )

summary(trackloss_summary)
head(avg_trackloss)
```

## 1.5. Reshape binomially encoded fixation data to a long format for plotting and analysis

```{r}
fix.long <- fix.dat.full %>%
  pivot_longer(
    cols = c(Target, Competitor, Distractor),
    names_to = "AOI",
    values_to = "FixP"
  )
```
