---
title: "Normalized_DPA"
output: html_document
date: "2026-05-30"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. Data preparation

```{r eval=T, echo=T, message=F}
rm(list = ls()) # Clear the workspace
options(scipen = 999) # Turn off scientific notation
```

## Packages

```{r eval=T, message=F}
require(Rmisc)
require(tidyverse)
require(ggplot2)
require(boot)
require(dplyr)
require(lme4)
require(purrr)
require(languageserver)
```

## Package versions

```{r}
installed.packages()[names(sessionInfo()$otherPkgs), "Version"]
```

## 1.1. Define constants

```{r}
# Window of interest (in milliseconds)
WINDOW_START <- -300
WINDOW_END <- 2000

# Number of pseudobins for normalization
N_PBINS = 100

# Number of consecutive statistically significant time bins required to establish divergence
Nbins = 5

# Number of bootstrap iterations
Niter = 100L
```

## 1.2. Read the time-course output data of the eye-tracker from file

```{r}
fix.dat = read_delim("data-2025-03-12/df_cleaned.csv", delim=',')
fix.dat = fix.dat %>% mutate_at(vars(Timeline), as.factor)
# all the trackloss bins have been removed - in any given row, there is a fixation in one of the 3 AreaFixated 

summary(fix.dat)
head(fix.dat)
```

## 1.3. Fill in the missing time points and add a trackloss column

```{r}
# Make a list of relevant trial variables
trial_vars <- c(
  "SpeakerType",
  "Determiner",
  "Phonol_Overlap_T",
  "NounOnset_adjusted",
  "EndingOffset_adjusted"
)

# Create rows for the missing time points
fix.dat.full <- fix.dat %>%
  
  filter(between(TimeMS_adjusted, WINDOW_START, WINDOW_END)) %>%
  
  group_by(Participant, TOI) %>%
  complete(
    TimeMS_adjusted = seq(
      WINDOW_START,
      WINDOW_END,
      by = 10
    )
  ) %>%
  
  # Fill in the values for the trial variables
  fill(all_of(trial_vars), .direction = "downup") %>%
  
  # Add binary fixation proportions of 0 to each AOI
  mutate(
    across(
      c(Target, Competitor, Distractor),
      ~ coalesce(.x, 0)
    )
  ) %>%
  ungroup()
```

## 1.4. Factorize relevant variables and drop extra columns

```{r}
fix.dat.full <- fix.dat.full %>%
  
  # Change independent variables to factors
  mutate(
    TOI = factor(TOI),
    Participant = factor(Participant),
    Determiner = factor(Determiner),
    SpeakerType = factor(SpeakerType),
    Phonol_Overlap_T = factor(Phonol_Overlap_T)
  ) %>%
  
  # Drop the extra columns
  select(
    Participant,
    TOI,
    TimeMS_adjusted,
    SpeakerType,
    Determiner,
    Phonol_Overlap_T,
    Target,
    Competitor,
    Distractor,
    NounOnset_adjusted,
    EndingOffset_adjusted
  ) %>%
  
  # Collapse the two types of non-minimal pairs in the phonological overlap variable
  mutate(
    Phonol_Overlap_T = forcats::fct_collapse(
      Phonol_Overlap_T,
      "Non Minimal Pair" = c(
        "First Syll-stressed",
        "First Syll-unstressed"
      )
    )
  ) %>%
  
  # Add an AreaFixated column
  mutate(
    AreaFixated = case_when(
      Target == 1 ~ "Target",
      Competitor == 1 ~ "Competitor",
      Distractor == 1 ~ "Distractor",
      TRUE ~ "Trackloss"
    )
  )

summary(fix.dat.full)
```

### Sanity check plot (step function plot for one participant and TOI)

```{r eval=F}
trial_id <- "Trial02-def-perro_0"
participant_id <- "TA105"

df_plot <- fix.dat.full %>%
  filter(Participant == participant_id, TOI == trial_id) %>%
  select(TimeMS_adjusted, Target, Competitor, Distractor) %>%
  pivot_longer(cols = c(Target, Competitor, Distractor),
               names_to = "AOI", values_to = "Fixation")

library(ggplot2)

vline_df <- tibble::tibble(
  time = c(108, 551),
  event = c("Noun onset", "Ending offset")
)

ggplot(df_plot, aes(x = TimeMS_adjusted, y = Fixation, color = AOI)) +
  geom_step(size = 0.8) +
  geom_vline(
    data = vline_df,
    aes(xintercept=time, linetype=event),
    color="black",
    linewidth=0.8
  ) +
  scale_x_continuous(limits = c(WINDOW_START, WINDOW_END)) +
  scale_y_continuous(limits = c(0, 1)) +
  labs(title = "Binary Fixations by AOI",
       subtitle = "Participant TA105 – Trial02-def-perro_0",
       x = "Time (ms)",
       y = "Fixated (0/1)") +
  theme_minimal()
```

### Average noun onset and ending offset

```{r eval=F}
fix.dat.full %>%
  summarise(avg_onset = mean(unique(NounOnset_adjusted), na.rm = TRUE))
fix.dat.full %>%
  summarise(avg_offset = mean(unique(EndingOffset_adjusted), na.rm = TRUE))

# Average noun onset: 127.6885 ms
# Average noun offset: 618.0482 ms
```

### Calculate the proportion of trackloss for each trial

(why are some trackloss proportions 1???)

```{r}
library(dplyr)

trackloss_summary <- fix.dat.full %>%
  
  # mark trackloss rows
  mutate(trackloss = AreaFixated == "Trackloss") %>%
  
  # classify time window
  mutate(window = case_when(
    TimeMS_adjusted >= 0 & TimeMS_adjusted < NounOnset_adjusted ~ "determiner_portion",
    TimeMS_adjusted >= NounOnset_adjusted & TimeMS_adjusted <= EndingOffset_adjusted ~ "noun_portion",
    TRUE ~ NA_character_
  )) %>%
  
  # keep only relevant time bins
  filter(!is.na(window)) %>%
  
  # compute proportion
  group_by(Participant, TOI, window) %>%
  summarise(
    prop_trackloss = mean(trackloss),
    n_samples = n(),
    .groups = "drop"
  )

avg_trackloss <- trackloss_summary %>%
  group_by(window) %>%
  summarise(
    avg_prop_trackloss = mean(prop_trackloss),
    .groups = "drop"
  )

summary(trackloss_summary)
head(avg_trackloss)
```

::: {style="color:red;"}
329 out of 3044 rows in trackloss_summary are 1.000 - which means that there is no data at all in that det/noun portion and the trial should be excluded

`sum(trackloss_summary$prop_trackloss > 0.99, na.rm = TRUE)`

\>\>\> 329
:::

# 2. Plotting before normalization

## 2.1. Preparing the data for visualization

```{r}
# Pivot the binomially encoded columns
fix.long <- fix.dat.full %>%
  pivot_longer(
    cols = c(Target, Competitor, Distractor),
    names_to = "AOI",
    values_to = "FixP"
  )

fix.long = fix.long %>% mutate_at(vars(AOI), as.factor)
```

## 2.2. Plots of mean fixation proportion

### Plot with all three AOIs, INCLUDING trackloss, over ALL conditions (no facets)

```{r, fig.width = 12}
# Calculate the mean fixation proportion for each AOI, in each time bin
fix.means = Rmisc::summarySE(fix.long, 
                             measurevar='FixP', 
                             groupvars=c('TimeMS_adjusted','AOI'),
                             na.rm=T)

# Make the plot
fix.plot.all = ggplot() + 
  geom_line(data=fix.means,aes(x=TimeMS_adjusted, y=FixP, group=AOI, colour=AOI, lty=AOI), lwd=1.5) +
  geom_ribbon(data=fix.means,aes(x=TimeMS_adjusted,ymin=FixP-se,ymax=FixP+se,color=AOI,fill=AOI), linewidth=.2, alpha=.3, lty="dashed", show.legend=F)  +
  labs(y="Fixation Proportion", x="Time relative to the target word onset (ms)") + 
  
  scale_x_continuous(limits=c(-300, 1400),expand=c(0,0),breaks=seq(-300, 1400, 200)) +
  scale_color_manual('AOI', values=c("pink","skyblue","gold")) +
  scale_fill_manual('AOI', values=c("pink","skyblue","gold")) +
  scale_linetype_manual('AOI', values=c("solid","dotted", "dotdash")) +
  
  theme(
    text=element_text(size=20),
    legend.key.height=unit(.3,"in"),
    legend.key.width=unit(.6,"in"),
    panel.spacing=unit(1, "in")) 

fix.plot.all
```

### Plot with all three AOIs, EXCLUDING trackloss, over ALL conditions (no facets)

```{r, fig.width = 12}

# Drop rows with trackloss
fix.dat.full.noloss = fix.dat.full %>% filter(AreaFixated %in% c('Target','Competitor', 'Distractor')) %>% droplevels()

# Pivot the binomially encoded columns
fix.long.noloss <- fix.dat.full.noloss %>%
  pivot_longer(
    cols = c(Target, Competitor, Distractor),
    names_to = "AOI",
    values_to = "FixP"
  )

fix.long.noloss = fix.long.noloss %>% mutate_at(vars(AOI), as.factor)

# Calculate the mean fixation proportion for each AOI, in each time bin
fix.means = Rmisc::summarySE(fix.long.noloss, 
                             measurevar='FixP', 
                             groupvars=c('TimeMS_adjusted','AOI'),
                             na.rm=T)

# Make the plot
fix.plot.all = ggplot() + 
  geom_line(data=fix.means,aes(x=TimeMS_adjusted, y=FixP, group=AOI, colour=AOI, lty=AOI), lwd=1.5) +
  geom_ribbon(data=fix.means,aes(x=TimeMS_adjusted,ymin=FixP-se,ymax=FixP+se,color=AOI,fill=AOI), linewidth=.2, alpha=.3, lty="dashed", show.legend=F)  +
  labs(y="Fixation Proportion", x="Time relative to the target word onset (ms)") + 
  
  scale_x_continuous(limits=c(-300, 1400),expand=c(0,0),breaks=seq(-300, 1400, 200)) +
  scale_color_manual('AOI', values=c("pink","skyblue","gold")) +
  scale_fill_manual('AOI', values=c("pink","skyblue","gold")) +
  scale_linetype_manual('AOI', values=c("solid","dotted", "dotdash")) +
  
  theme(
    text=element_text(size=20),
    legend.key.height=unit(.3,"in"),
    legend.key.width=unit(.6,"in"),
    panel.spacing=unit(1, "in")) 

fix.plot.all
```

# 3. Normalization

## 3.1 Create one step function per Participant × AOI × TOI

```{r}
# Make sure FixP is not a factor
fix.long <- fix.long %>%
  mutate(FixP = as.numeric(as.character(FixP)))

step_functions <- fix.long %>%
  group_by(Participant, AOI, TOI) %>%
  summarise(
    f = list(approxfun(
      x = TimeMS_adjusted,
      y = FixP,
      method = "constant",
      rule = 2
    )),
    .groups = "drop"
  )
```

### Plot the step functions for one participant-trial

```{r eval=F}
step_functions_example = step_functions %>% filter(Participant == "TA105", TOI == "Trial02-def-perro_0") 

time_grid <- seq(WINDOW_START, WINDOW_END, by = 10)

library(purrr)

plot_df <- step_functions_example %>%
  mutate(
    data = map(f, ~ tibble(
      TimeMS_adjusted = time_grid,
      FixP = .x(time_grid)
    ))
  ) %>%
  unnest(data)

library(ggplot2)

ggplot(plot_df, aes(TimeMS_adjusted, FixP, colour = AOI)) +
  geom_step(linewidth = 1) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  theme_bw() +
  labs(
    x = "Time (ms)",
    y = "FixP",
    title = "Step function: TA105 Trial02-def-perro_0"
  )
```

## 3.2 Normalization for determiner portion

### 3.2.1 Define windows for each trial

```{r eval=T, message=F}
# Compute trial-specific windows
trial_windows_det <- fix.long %>%
  group_by(TOI) %>%
  summarise(
    SpeakerType = first(SpeakerType),
    Determiner  = first(Determiner),
    Phonol_Overlap_T = first(Phonol_Overlap_T),
    start_anchor = 0,
    end_anchor   = max(NounOnset_adjusted, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    window_list = map2(start_anchor, end_anchor, function(start, end) {
      edges <- seq(start, end, length.out = N_PBINS + 1)
      tibble(start = head(edges, -1), end = tail(edges, -1))
    })
  ) %>%
  unnest(window_list)
```

### 3.2.2 Compute AUCs for all step functions

```{r}
auc_trial_windows_det <- step_functions %>%
  left_join(trial_windows_det, by = "TOI", relationship = "many-to-many") %>%
  mutate(
    AUC = pmap_dbl(
      list(f, start, end),
      function(fun, start, end) {
        tryCatch(
          integrate(fun, lower = start, upper = end)$value,
          error = function(e) NA_real_
        )
      }
    ),
    WindowDur = end - start,
    AUC_prop = AUC / WindowDur
  )
```

### 3.2.3 Add pseudo-bin numbers

```{r eval=T, message=F}
auc_trial_windows_det <- auc_trial_windows_det %>%
  group_by(Participant, TOI, SpeakerType, Determiner, Phonol_Overlap_T) %>%
  arrange(start, AOI, .by_group = TRUE) %>%
  mutate(PBin = rep(1:N_PBINS, each = length(unique(AOI)), length.out = n())) %>%
  ungroup()
```

### 3.2.4 Compute average AUCs across conditions (for smooth plotting)

```{r eval=T, message=F}
auc_summary_det <- auc_trial_windows_det %>%
  group_by(AOI, PBin, SpeakerType, Determiner, Phonol_Overlap_T) %>%
  summarise(
    n = n(),
    mean_AUC = mean(AUC, na.rm = TRUE),
    se_AUC = sd(AUC, na.rm = TRUE) / sqrt(n),
    mean_AUC_prop = mean(AUC_prop, na.rm = TRUE),
    se_AUC_prop = sd(AUC_prop, na.rm = TRUE) / sqrt(n),
    .groups = "drop"
  )
```

### 3.2.5 Plot the summary of AUC proportions across trials and participants

```{r eval=F}
ggplot(
  auc_summary_det,
  aes(
    x = PBin,
    y = mean_AUC_prop,
    colour = AOI,
    group = AOI
  )
) +
  geom_line(linewidth = 1) +
  geom_ribbon(
    aes(
      ymin = mean_AUC_prop - se_AUC_prop,
      ymax = mean_AUC_prop + se_AUC_prop,
      fill = AOI
    ),
    alpha = 0.2,
    colour = NA
  ) +
  facet_grid(
    SpeakerType ~ Determiner + Phonol_Overlap_T
  ) +
  scale_color_manual(
    values = c(
      "Target" = "gold",
      "Competitor" = "pink",
      "Distractor" = "skyblue"
    )
  ) +
  scale_fill_manual(
    values = c(
      "Target" = "gold",
      "Competitor" = "pink",
      "Distractor" = "skyblue"
    )
  ) +
  scale_x_continuous(
    breaks = seq(0, 100, by = 20)
  ) +
  scale_y_continuous(
    limits = c(0, 1)
  ) +
  labs(
    title = "Mean fixation proportions across normalized determiner pseudobins",
    x = "Normalized determiner pseudobin",
    y = "Mean fixation proportion",
    colour = "AOI",
    fill = "AOI"
  ) +
  theme_minimal(base_size = 16) +
  theme(
    strip.text = element_text(face = "bold"),
    legend.position = "bottom"
  )
```

## 3.3 Normalization for noun portion

### 3.3.1 Define windows for each trial

```{r eval=T, message=F}
# Compute trial-specific windows
trial_windows_noun <- fix.long %>%
  group_by(TOI) %>%
  summarise(
    SpeakerType = first(SpeakerType),
    Determiner = first(Determiner),
    Phonol_Overlap_T = first(Phonol_Overlap_T),
    start_anchor = min(NounOnset_adjusted, na.rm = TRUE),
    end_anchor = max(EndingOffset_adjusted, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    window_list = map2(start_anchor, end_anchor, function(start, end) {
      edges <- seq(start, end, length.out = N_PBINS + 1)
      tibble(
        start = head(edges, -1),
        end = tail(edges, -1)
      )
    })
  ) %>%
  unnest(window_list)
```

### 3.3.2 Compute AUCs for all step functions

```{r}
auc_trial_windows_noun <- step_functions %>%
  left_join(
    trial_windows_noun,
    by = "TOI",
    relationship = "many-to-many"
  ) %>%
  mutate(
    AUC = pmap_dbl(
      list(f, start, end),
      function(fun, start, end) {
        tryCatch(
          integrate(fun, lower = start, upper = end)$value,
          error = function(e) NA_real_
        )
      }
    ),
    WindowDur = end - start,
    AUC_prop = AUC / WindowDur
  )
```

### 3.3.3 Add pseudo-bin numbers

```{r eval=T, message=F}
auc_trial_windows_noun <- auc_trial_windows_noun %>%
  group_by(
    Participant,
    TOI,
    SpeakerType,
    Determiner,
    Phonol_Overlap_T
  ) %>%
  arrange(start, AOI, .by_group = TRUE) %>%
  mutate(
    PBin = rep(
      1:N_PBINS,
      each = length(unique(AOI)),
      length.out = n()
    ) + N_PBINS
  ) %>%
  ungroup()
```

### 3.3.4 Compute average AUCs across conditions (for smooth plotting)

```{r eval=T, message=F}
auc_summary_noun <- auc_trial_windows_noun %>%
  group_by(
    AOI,
    PBin,
    SpeakerType,
    Determiner,
    Phonol_Overlap_T
  ) %>%
  summarise(
    n = n(),
    mean_AUC = mean(AUC, na.rm = TRUE),
    se_AUC = sd(AUC, na.rm = TRUE) / sqrt(n),
    mean_AUC_prop = mean(AUC_prop, na.rm = TRUE),
    se_AUC_prop = sd(AUC_prop, na.rm = TRUE) / sqrt(n),
    .groups = "drop"
  )
```

### 3.3.5 Plot the summary of AUC proportions across trials and participants

```{r eval=F}
ggplot(
  auc_summary_noun,
  aes(
    x = PBin,
    y = mean_AUC_prop,
    colour = AOI,
    group = AOI
  )
) +
  geom_line(linewidth = 1) +
  geom_ribbon(
    aes(
      ymin = mean_AUC_prop - se_AUC_prop,
      ymax = mean_AUC_prop + se_AUC_prop,
      fill = AOI
    ),
    alpha = 0.2,
    colour = NA
  ) +
  facet_grid(
    SpeakerType ~ Determiner + Phonol_Overlap_T
  ) +
  scale_color_manual(
    values = c(
      "Target" = "gold",
      "Competitor" = "pink",
      "Distractor" = "skyblue"
    )
  ) +
  scale_fill_manual(
    values = c(
      "Target" = "gold",
      "Competitor" = "pink",
      "Distractor" = "skyblue"
    )
  ) +
  scale_x_continuous(
    breaks = seq(100, 200, by = 20)
  ) +
  scale_y_continuous(
    limits = c(0, 1)
  ) +
  labs(
    title = "Mean fixation proportions across normalized noun pseudobins",
    x = "Normalized noun pseudobin",
    y = "Mean fixation proportion",
    colour = "AOI",
    fill = "AOI"
  ) +
  theme_minimal(base_size = 16) +
  theme(
    strip.text = element_text(face = "bold"),
    legend.position = "bottom"
  )
```

## 3.4 Plotting after normalization

### 3.4.1 Calculate mean fixation for the remainder, 700ms after each trial's ending offset

```{r}
fix.long.post <- fix.long %>%
  group_by(Participant, TOI) %>%
  
  # keep only the 0–700 ms window after ending offset
  filter(
    TimeMS_adjusted >= EndingOffset_adjusted,
    TimeMS_adjusted <= EndingOffset_adjusted + 700
  ) %>%
  
  mutate(
    PBin = ((TimeMS_adjusted - min(TimeMS_adjusted)) / 10) + 1 + N_PBINS*2
  ) %>%
  
  ungroup()

# fix.long.notrackloss <- fix.long.post %>%
#   dplyr::filter(!Trackloss)

fix.means.long = Rmisc::summarySE(fix.long.post, measurevar='FixP', groupvars=c('PBin','SpeakerType', 'AOI', 'Determiner', 'Phonol_Overlap_T'), na.rm=T)

fix.means.long <- fix.means.long %>%
  rename(mean_AUC_prop = FixP)

summary(fix.means.long)
head(fix.means.long)
```

### 3.4.2 Concatenated plots with normalized determiner portions and noun portions, and non-normalized remainders

```{r}
# Combine determiner + noun windows
auc_combined <- bind_rows(
  auc_summary_det,
  auc_summary_noun,
  fix.means.long
)

# Make sure grouping vars are factors
auc_combined <- auc_combined %>%
  mutate(
    SpeakerType = as.factor(SpeakerType),
    Determiner = as.factor(Determiner),
    Phonol_Overlap_T = as.factor(Phonol_Overlap_T)
  )

# Plot
auc_plot <- ggplot(
  auc_combined,
  aes(
    x = PBin,
    y = mean_AUC_prop,
    color = AOI,
    group = AOI
  )
) +

  # 8 panels now
  facet_wrap(
    ~ Determiner + SpeakerType + Phonol_Overlap_T,
    ncol = 4,
    labeller = labeller(
      Determiner = c(
        "Definite" = "Definite determiner",
        "Possessive" = "Possessive determiner"
      ),
      SpeakerType = c(
        "1" = "ASI",
        "2" = "HSS"
      ),
      Phonol_Overlap_T = c(
        "Minimal Pair" = "Minimal Pair",
        "NonMinPair" = "Non Minimal Pair"
      )
    )
  ) +

  theme_bw() +

  # Line
  geom_line(linewidth = 1.2) +

  # SE ribbon
  geom_ribbon(
    aes(
      ymin = mean_AUC_prop - se_AUC_prop,
      ymax = mean_AUC_prop + se_AUC_prop,
      fill = AOI
    ),
    alpha = 0.2,
    color = NA
  ) +

  # Vertical line at noun onset
  geom_vline(
    xintercept = N_PBINS,
    linetype = "dashed",
    linewidth = 0.8
  ) +
  
  # Vertical line at noun offset
  geom_vline(
    xintercept = N_PBINS * 2,
    linetype = "dashed",
    linewidth = 0.8
  ) +
  
  annotate(
    "text",
    x = N_PBINS,
    y = 0.72,
    label = "Noun onset",
    angle = 90,
    vjust = -0.5,
    size = 5
  ) +
  
  annotate(
    "text",
    x = N_PBINS * 2,
    y = 0.72,
    label = "Noun offset",
    angle = 90,
    vjust = -0.5,
    size = 5
  ) +
  
  # Horizontal reference lines
  geom_hline(
    yintercept = c(0.33, 0.67),
    color = "gray75",
    linewidth = 0.7,
    linetype = "solid"
  ) +

  scale_color_manual(
    values = c(
      "Target" = "gold",
      "Competitor" = "pink",
      "Distractor" = "skyblue"
    )
  ) +

  scale_fill_manual(
    values = c(
      "Target" = "gold",
      "Competitor" = "pink",
      "Distractor" = "skyblue"
    )
  ) +

  labs(
    title = "Mean Fixation Proportion over Time: Determiner → Noun → Remainder",
    x = "Pseudo-bin (normalized time)",
    y = "Fixation proportion",
    color = "AOI",
    fill = "AOI"
  ) +

  theme(
    text = element_text(size = 18),
    panel.spacing = unit(1, "cm"),
    legend.position = "right"
  )

auc_plot
```

```{r}
# Combine determiner + noun windows
auc_combined <- bind_rows(
  auc_summary_det,
  auc_summary_noun,
  fix.means.long
)

# Aggregate across Minimal Pair vs Non-Minimal Pair
auc_combined_4facet <- auc_combined %>%
  group_by(
    AOI,
    PBin,
    SpeakerType,
    Determiner
  ) %>%
  summarise(
    mean_AUC_prop = mean(mean_AUC_prop, na.rm = TRUE),
    
    # Approximate combined SE
    se_AUC_prop = sqrt(
      sum(se_AUC_prop^2, na.rm = TRUE)
    ) / sum(!is.na(se_AUC_prop)),
    
    .groups = "drop"
  ) %>%
  mutate(
    SpeakerType = as.factor(SpeakerType),
    Determiner = as.factor(Determiner)
  )


# Plot
auc_plot <- ggplot(
  auc_combined_4facet,
  aes(
    x = PBin,
    y = mean_AUC_prop,
    color = AOI,
    group = AOI
  )
) +

  # 4 panels: Determiner × Speaker Type
  facet_wrap(
    ~ Determiner + SpeakerType,
    ncol = 2,
    labeller = labeller(
      Determiner = c(
        "Definite" = "Definite determiner",
        "Possessive" = "Possessive determiner"
      ),
      SpeakerType = c(
        "1" = "ASI",
        "2" = "HSS"
      )
    )
  ) +

  theme_bw() +

  # Line
  geom_line(linewidth = 1.2) +

  # SE ribbon
  geom_ribbon(
    aes(
      ymin = mean_AUC_prop - se_AUC_prop,
      ymax = mean_AUC_prop + se_AUC_prop,
      fill = AOI
    ),
    alpha = 0.2,
    color = NA
  ) +

  # Noun onset
  geom_vline(
    xintercept = N_PBINS,
    linetype = "dashed",
    linewidth = 0.8
  ) +

  # Noun offset
  geom_vline(
    xintercept = N_PBINS * 2,
    linetype = "dashed",
    linewidth = 0.8
  ) +

  annotate(
    "text",
    x = N_PBINS,
    y = 0.72,
    label = "Noun onset",
    angle = 90,
    vjust = -0.5,
    size = 5
  ) +

  annotate(
    "text",
    x = N_PBINS * 2,
    y = 0.72,
    label = "Noun offset",
    angle = 90,
    vjust = -0.5,
    size = 5
  ) +

  # Horizontal reference lines
  geom_hline(
    yintercept = c(0.33, 0.67),
    color = "gray75",
    linewidth = 0.7
  ) +

  scale_color_manual(
    values = c(
      "Target" = "gold",
      "Competitor" = "pink",
      "Distractor" = "skyblue"
    )
  ) +

  scale_fill_manual(
    values = c(
      "Target" = "gold",
      "Competitor" = "pink",
      "Distractor" = "skyblue"
    )
  ) +

  labs(
    title = "Mean Fixation Proportion over Time",
    x = "Pseudo-bin (normalized time)",
    y = "Fixation proportion",
    color = "AOI",
    fill = "AOI"
  ) +

  theme(
    text = element_text(size = 18),
    panel.spacing = unit(1, "cm"),
    legend.position = "right"
  )

auc_plot
```

# 4. DPA

Here we will run a by-participant analysis testing the divergence between the Target and NonTarget AOIs. The conditions are SpeakerType and Determiner.

## 4.1 Preparations

### 4.1.1 Concatenate the normalized time series for determiners and nouns *before* taking the average across participants and trials

We cannot use `auc_combined` because it takes the average of FixP over all participants and trials

```{r}
auc_combined_raw <- bind_rows(auc_trial_windows_det, auc_trial_windows_noun, fix.long.post)

auc_combined_raw <- auc_combined_raw %>% select(-f, -TimeMS_adjusted, -FixP) 
```

### \* Empirical logit

```{r}
auc_combined_raw = auc_combined_raw %>% mutate(logFix = log(AUC_prop/(1-AUC_prop)))

Nbins = 5
```

::: {style="color: red;"}
### Aug 14 edit - new DPA groups

#### Filter for alternating nouns only in the HSS group

```{r}
div.dat.asi <- auc_combined_raw %>% filter(SpeakerType == "1")
div.dat.hss <- auc_combined_raw %>% filter(SpeakerType == "2" & Phonol_Overlap_T == "Minimal Pair")
```

#### Reduce the dataset to target only

```{r}
div.dat.asi <- div.dat.asi %>% filter(AOI == "Target")
div.dat.hss <- div.dat.hss %>% filter(AOI == "Target")
```

#### Create new stratification variables

```{r}
div.dat.asi = div.dat.asi %>%  
  filter(PBin<200) %>%
  dplyr::mutate(Determiner = relevel(as.factor(Determiner), ref = "Definite"),
                StrataVars=paste(Participant, Determiner, PBin, sep='')) %>% 
  mutate_at(vars(StrataVars),as.factor) %>% 
  droplevels()

div.dat.hss = div.dat.hss %>%  
  filter(PBin<200) %>%
  dplyr::mutate(Determiner = relevel(as.factor(Determiner), ref = "Definite"),
                StrataVars=paste(Participant, Determiner, PBin, sep='')) %>% 
  mutate_at(vars(StrataVars),as.factor) %>% 
  droplevels()
```

#### Rename columns

```{r}
div.dat.asi <- div.dat.asi %>%
  rename(
    TimeMS_adjusted = PBin,
    FixP = AUC_prop
  )

div.dat.hss <- div.dat.hss %>%
  rename(
    TimeMS_adjusted = PBin,
    FixP = AUC_prop
  )
```
:::

### 4.1.2 Reduce the dataset to two variables

```{r}
nontarget <- auc_combined_raw %>%
  filter(AOI %in% c("Competitor")) %>%
  group_by(
    Participant,
    TOI,
    PBin,
    SpeakerType,
    Determiner,
    Phonol_Overlap_T
  ) %>%
  summarise(
    AUC_prop = sum(AUC_prop, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(AOI = "NonTarget")

target <- auc_combined_raw %>%
  filter(AOI == "Target")

auc_combined_raw_twovar <- bind_rows(
  target,
  nontarget
)
```

### 4.1.3 Create stratification variables

Previously we selected a time window; now it will be the normalized and concatenated determiner and noun portions, as well as the non-normalized remainder. Now the pseudo-bin numbers will act like the previous time bin numbers `TimeMS_adjusted` ; FixP will become AUC_prop.

```{r eval=T, echo=T, message=F}
div.dat = auc_combined_raw_twovar %>%  
  filter(PBin<200) %>%
  dplyr::mutate(StrataVars=paste(Participant, AOI, PBin, sep='')) %>% 
  mutate_at(vars(SpeakerType, StrataVars),as.factor) %>% 
  droplevels()

summary(div.dat)
head(div.dat)
```

### 4.1.4 Rename columns so we don't have to change the original DPA code

```{r}
div.dat <- div.dat %>%
  rename(
    TimeMS_adjusted = PBin,
    FixP = AUC_prop
  )
```

### 4.1.5 Create two subsets based on the type of determiner

```{r eval=T, echo=T, message=F}
div.dat.def <- div.dat %>% filter(Determiner == "Definite")
div.dat.pos <- div.dat %>% filter(Determiner == "Possessive")
```

## 4.2 Define bootstrap functions to compute a divergence point for the target vs. unrelated conditions

:::: {style="color:red;"}
### Aug 14 edit - wrote new bootstrap function using lmer

```{r eval=T, echo=T, message=F}

boot_determiner_lmer <- function(original_data, resample_indices) {

  dat <- original_data[resample_indices, ]

  test <- dat %>%
    group_by(TimeMS_adjusted) %>%
    group_modify(~ {

      fit <- lmer(
        FixP ~ Determiner + (1 | Participant),
        data = .x
      )

      tibble(
        t = summary(fit)$coefficients[
          "DeterminerPossessive",
          "t value"
        ]
      )
    }) %>%
    ungroup()

  # Crystal said "when Definite and Possessive DIFFER"
  # so use a two-sided threshold
  significant <- abs(test$t) > 1.96

  # Find earliest run of Nbins consecutive significant bins
  onset <- rep(FALSE, length(significant))

  if (length(significant) >= Nbins) {

    for (i in 1:(length(significant) - Nbins + 1)) {

      onset[i] <- all(
        significant[i:(i + Nbins - 1)],
        na.rm = FALSE
      )

    }
  }

  onset_index <- which(onset)[1]

  if (length(onset_index) == 0) {
    return(NA_real_)
  }

  # Return the ACTUAL time rather than merely the index
  test$TimeMS_adjusted[onset_index]
}

# boot_L1L2_lmer = function(original_data, resample_indices){
#   dat = original_data[resample_indices, ]
#   
#   # a statistical test at each timepoint for each group
# #  test_g1 = dat %>% # t-test for L1 group
# #    subset(SpeakerType == 1) %>% group_by(TimeMS_adjusted) %>%
# #    dplyr::summarise(t = summary(lmer(FixP ~ AOI + (1|Participant)))$coefficients["AOITarget", "t value"])
#   
#   test_g1 <- tryCatch(
# 
#     dat %>%
#         subset(Determiner=="Definite") %>%
#         group_by(TimeMS_adjusted) %>%
#         group_modify(~{
# 
#           fit <- lmer(
#               FixP~AOI+
#               (1|Participant),
#               data=.x
#           )
#       
#           tibble(
#       
#               t=
#                   summary(fit)$coefficients[
#                       "AOITarget",
#                       "t value"
#                   ]
#       
#           )
#       
#       }),
# 
#     error=function(e){
# 
#         cat("\nMODEL FAILED\n")
# 
#         print(
#             dat %>%
#             subset(SpeakerType==1) %>%
#             filter(TimeMS_adjusted==201) %>%
#             count(AOI)
#         )
# 
#         print(
#             dat %>%
#             subset(SpeakerType==1) %>%
#             filter(TimeMS_adjusted==201) %>%
#             count(Participant)
#         )
# 
#         stop(e)
# 
#     }
# 
# )
# 
#   test_g2 = dat %>% # t-test for L2 group
#     subset(Determiner == "Possessive") %>% group_by(TimeMS_adjusted) %>%
#     dplyr::summarise(t = summary(lmer(FixP ~ AOI + (1|Participant)))$coefficients["AOITarget", "t value"])
# 
#   # return a TRUE/FALSE vector of significant positive t-scores
#   # (positive means more looks to the target than unrelated)
#   t_g1 = test_g1$t > 1.96
#   t_g2 = test_g2$t > 1.96
#   
#   # create empty vectors to store onsets
#   onset_g1 = onset_g2 = c()
#   
#   # find the index of the earliest run of sequential TRUEs; threshold is Nbins
#   for (i in 1:(length(t_g2)-Nbins)) {
#     onset_g1[i] = sum(t_g1[i:(i+Nbins-1)]) == Nbins
#     onset_g2[i] = sum(t_g2[i:(i+Nbins-1)]) == Nbins
#   }
#   
#   # find the difference between onsets
#   delta_g1g2 = which(onset_g2)[1] - which(onset_g1)[1]
#   
#   # print 
#   # note: the bootstrap returns the indices of the respective timepoints, not absolute times. 
#   # The annotations to the right of each index (e.g. t[,1]) indicate where in the boot object the bootstrapped onset distributions can be found.
#   c(delta_g1g2,         # onset difference L1 vs. L2 t[,1]
#     which(onset_g1)[1], # onset bin for looks to target L1 t[,2]
#     which(onset_g2)[1])  # onset bin for looks to target L2 t[,3]
# }

```

### Run the new bootstrap function - returns "boundary (singular) fit"

::: {style="color:red;"}
```{r}
bootres_determiner_lmer_asi = boot::boot(
  data = div.dat.asi,  # data set to bootstrap       
  statistic = boot_determiner_lmer,  # bootstrap function      
  strata = div.dat.asi$StrataVars, # stratification variable 
  R = Niter)  # number of iterations

bootres_determiner_lmer_hss = boot::boot(
  data = div.dat.hss,  # data set to bootstrap       
  statistic = boot_determiner_lmer,  # bootstrap function      
  strata = div.dat.hss$StrataVars, # stratification variable 
  R = Niter)  # number of iterations

bootres_determiner_lmer_asi
bootres_determiner_lmer_hss

save(bootres_determiner_lmer_asi, file="NEW.DPA_results.lmer.asi.rds")
save(bootres_determiner_lmer_hss, file="NEW.DPA_results.lmer.hss.rds")
```
:::
::::

### 4.2.2 Run the lmer bootstrap on the two subsets by determiner type

```{r eval=F, echo=T, message=F}
bootres_L1L2_lmer_def = boot::boot(
  data = div.dat.def,  # data set to bootstrap       
  statistic = boot_L1L2_lmer,  # bootstrap function      
  strata = div.dat.def$StrataVars, # stratification variable 
  R = Niter)  # number of iterations

bootres_L1L2_lmer_pos = boot::boot(
  data = div.dat.pos,  # data set to bootstrap       
  statistic = boot_L1L2_lmer,  # bootstrap function      
  strata = div.dat.pos$StrataVars, # stratification variable 
  R = Niter)  # number of iterations

bootres_L1L2_lmer_def
bootres_L1L2_lmer_pos
```

Save the lmer results

```{r eval=F, echo=T, message=F}
save(bootres_L1L2_lmer_def, file="NEW.DPA_results.lmer.def.rds")
save(bootres_L1L2_lmer_pos, file="NEW.DPA_results.lmer.pos.rds")
```

Define a function that runs the bootstrap function for all the trials

```{r eval=F, echo=T, message=F}
run_boot_per_trial <- function(toi_id, df, Niter = 100, strata_col = "StrataVars") {
  trial_data <- df %>% filter(TOI == toi_id) %>% droplevels()

  boot_result <- boot::boot(
    data = trial_data,
    statistic = boot_L1L2_glm,  # your divergence function
    strata = trial_data[[strata_col]],
    R = Niter
  )

  # Extract metadata from the first row (assuming Determiner etc. are constant per trial)
  trial_info <- trial_data %>%
    slice(1) %>%
    select(TOI, Determiner, Phonol_Overlap, GenderMatch)  # add more if needed

  list(boot = boot_result, info = trial_info)
}
```

Run the function that runs the bootstrap function

```{r eval=F, echo=T, message=F}
trial_ids <- unique(div.dat$TOI)

# Store both the bootstrap and metadata
boot_trials <- lapply(trial_ids, function(toi) run_boot_per_trial(toi, div.dat, Niter = 100))
names(boot_trials) <- trial_id
```

Extract divergence values + metadata into a summary dataframe

```{r eval=F, echo=T, message=F}
# Function to extract means and metadata
extract_div_summary <- function(result) {
  bootres <- result$boot
  info <- result$info

  tibble::tibble(
    TOI = info$TOI,
    Determiner = info$Determiner,
    Phonol_Overlap = info$Phonol_Overlap,
    GenderMatch = info$GenderMatchTarget_D,
    ASI_divergence = mean(bootres$t[,2], na.rm = TRUE) * 10 - 10,
    HSS_divergence = mean(bootres$t[,3], na.rm = TRUE) * 10 - 10
  )
}

# Build summary
div_summary_df <- purrr::map_dfr(boot_trials, extract_div_summary)
```

Bootstrap function using glm (generalized linear model)

```{r eval=F, echo=T, message=F}
boot_L1L2_glm = function(original_data, resample_indices){
  dat = original_data[resample_indices, ]
  
  # a statistical test at each timepoint for each group
  test_g1 = dat %>% # t-test for L1 group
    subset(SpeakerType == 1) %>% group_by(TimeMS_adjusted) %>%
    dplyr::summarise(t = summary(glm(FixP ~ AOI, family = binomial))$coefficients["AOITarget", "z value"])

  test_g2 = dat %>% # t-test for L2 group
    subset(SpeakerType == 2) %>% group_by(TimeMS_adjusted) %>%
    dplyr::summarise(t = summary(glm(FixP ~ AOI, family = binomial))$coefficients["AOITarget", "z value"])

  # return a TRUE/FALSE vector of significant positive t-scores
  # (positive means more looks to the target than unrelated)
  t_g1 = test_g1$t > 1.96
  t_g2 = test_g2$t > 1.96
  
  # create empty vectors to store onsets
  onset_g1 = onset_g2 = c()
  
  # find the index of the earliest run of sequential TRUEs; threshold is Nbins
  for (i in 1:(length(t_g2)-Nbins)) {
    onset_g1[i] = sum(t_g1[i:(i+Nbins-1)]) == Nbins
    onset_g2[i] = sum(t_g2[i:(i+Nbins-1)]) == Nbins
  }
  
  # find the difference between onsets
  delta_g1g2 = which(onset_g2)[1] - which(onset_g1)[1]
  
  # print 
  # note: the bootstrap returns the indices of the respective timepoints, not absolute times. 
  # The annotations to the right of each index (e.g. t[,1]) indicate where in the boot object the bootstrapped onset distributions can be found.
  c(delta_g1g2,         # onset difference L1 vs. L2 t[,1]
    which(onset_g1)[1], # onset bin for looks to target L1 t[,2]
    which(onset_g2)[1])  # onset bin for looks to target L2 t[,3]
}

```

Run the glm bootstrap on the two subsets by determiner type

```{r eval=F, echo=T, message=F}
bootres_L1L2_glm_def = boot::boot(
  data = div.dat.def,  # data set to bootstrap       
  statistic = boot_L1L2_glm,  # bootstrap function      
  strata = div.dat.def$StrataVars, # stratification variable 
  R = Niter)  # number of iterations

bootres_L1L2_glm_pos = boot::boot(
  data = div.dat.pos,  # data set to bootstrap       
  statistic = boot_L1L2_glm,  # bootstrap function      
  strata = div.dat.pos$StrataVars, # stratification variable 
  R = Niter)  # number of iterations

bootres_L1L2_glm_def
bootres_L1L2_glm_pos
```

Save the glm results

```{r eval=F, echo=T, message=F}
save(bootres_L1L2_glm_def, file="DPA_results.glm.def.rds")
save(bootres_L1L2_glm_pos, file="DPA_results.glm.pos.rds")
```

##### Extracting and plotting the divergence points

Load the lmer results

```{r eval=T, echo=T, message=F}
load("DPA_results.lmer.def.rds")
load("DPA_results.lmer.pos.rds")
```

Load the glm results

```{r eval=F, echo=T, message=F}
load("DPA_results.glm.def.rds")
load("DPA_results.glm.pos.rds")
```

Define function to extract divergence points and confidence intervals from bootstrapping results

```{r eval=T, echo=T, message=F}
get_divergence_ms <- function(bootres, index) {
  # index = 2 for ASI, 3 for HSS
  mean_bin <- mean(bootres$t[, index], na.rm = TRUE)
  ci_bin <- boot::boot.ci(bootres, index = index, type = "perc")$percent[4:5]
  
  mean_ms <- (mean_bin - 1) * 10
  ci_ms <- (ci_bin - 1) * 10
  
  list(mean = mean_ms, ci = ci_ms)
}
```

Run the function on all the results

```{r eval=T, echo=T, message=F}
# Definite determiner
def_asi <- get_divergence_ms(bootres_L1L2_lmer_def, index = 2)
def_hss <- get_divergence_ms(bootres_L1L2_lmer_def, index = 3)

# Possessive determiner
pos_asi <- get_divergence_ms(bootres_L1L2_lmer_pos, index = 2)
pos_hss <- get_divergence_ms(bootres_L1L2_lmer_pos, index = 3)
```

Create dataframe for plotting

```{r eval=T, echo=T, message=F}
div_df <- tibble::tibble(
  Determiner = c("Definite", "Definite", "Possessive", "Possessive"),
  SpeakerType = c("1", "2", "1", "2"),
  TimeMS = c(def_asi$mean, def_hss$mean, pos_asi$mean, pos_hss$mean),
  CI_low = c(def_asi$ci[1], def_hss$ci[1], pos_asi$ci[1], pos_hss$ci[1]),
  CI_high = c(def_asi$ci[2], def_hss$ci[2], pos_asi$ci[2], pos_hss$ci[2])
)

div_df = div_df %>% mutate_at(vars(SpeakerType, Determiner), as.factor)

div_df <- div_df %>%
  mutate(label = paste0(TimeMS, " ms"))
```

Add divergence points and their confidence intervals to the two-variable plots

```{r eval=T, echo=T, message=F, fig.width = 14}
fix.plot.twovar +
  geom_point(data = div_df,
             aes(x = TimeMS, y = 0.48),
             shape = 18, size = 4, color = "black") +
  geom_errorbarh(data = div_df,
                 aes(xmin = CI_low, xmax = CI_high, y = 0.48),
                 height = 0.02, color = "black", linewidth = 0.6) +
  geom_text(data = div_df,
          aes(x = TimeMS - 50, y = 0.55, label = label), 
          size = 10, fontface="bold")
```

Save the plots

```{r eval=T, message=F}
ggsave("plots/fixation_plot_twovar_divpts.png", width = 20, height = 15, dpi = 300)
```

## 5. Group comparison

We can compute a p-value for the group comparison by creating a bootstrap distribution of the null hypothesis. To do this, we will first pool the original data from both groups, randomly assign group labels, and then estimate a difference in divergence point. We will repeat this many times to obtain a distribution of divergence point that could be expected under the null hypothesis. We will obtain the p-value by calculating the proportion of samples from this null distribution that are larger than the observed difference in divergence point in the empirical data.

Below, we will calculate a p-value for the group comparison (ASI vs. HSS group). We will first define a bootstrap function.

##### Define the bootstrap function for group comparison (only lmer so far)

```{r eval=T, echo=T, message=F}
boot_L1L2_pval = function(original_data, resample_indices){
  dat_resample = original_data[resample_indices, ] # resample the data
  
  dat = dat_resample %>% 
    group_by(Participant, AOI, TimeMS_adjusted) %>% 
    transform(SpeakerType=sample(SpeakerType,replace=F)) %>% # randomly assign group labels 
    ungroup() #%>% 
  
  # a statistical test at each timepoint for each group
  test_g1 = dat %>% # t-test for L1 group
    subset(SpeakerType == 1) %>% group_by(TimeMS_adjusted) %>%
    dplyr::summarise(t = summary(lmer(FixP ~ AOI + (1|Participant)))$coefficients["AOITarget", "t value"])

  test_g2 = dat %>% # t-test for L2 group
    subset(SpeakerType == 2) %>% group_by(TimeMS_adjusted) %>%
    dplyr::summarise(t = summary(lmer(FixP ~ AOI + (1|Participant)))$coefficients["AOITarget", "t value"])
  
  # return a TRUE/FALSE vector of significant positive t-scores  
  # (positive means more looks to the target than unrelated)
  t_g1 = test_g1$t > 1.96
  t_g2 = test_g2$t > 1.96
  
  # create empty vectors to store onsets
  onset_g1 = onset_g2 = c()
  
  # find the index of the earliest run of 4 sequential TRUEs 
  for (i in 1:(length(t_g2)-Nbins)) { 
    onset_g1[i] = sum(t_g1[i:(i+Nbins-1)]) == Nbins
    onset_g2[i] = sum(t_g2[i:(i+Nbins-1)]) == Nbins
  }
  
  # find the difference between onsets
  delta_g1g2 = which(onset_g2)[1] - which(onset_g1)[1]
  
  # print 
  delta_g1g2  # onset difference L1 vs. L2 t[,1]
}

```

##### Run the lmer bootstrap on the two subsets by determiner type

```{r eval=F, echo=T, message=F}
bootres_L1L2_pval_def = boot::boot(
  data = div.dat.def,  # data set to bootstrap       
  statistic = boot_L1L2_pval,  # bootstrap function      
  strata = div.dat.def$StrataVars, # stratification variable 
  R = Niter)  # number of iterations        

bootres_L1L2_pval_pos = boot::boot(
  data = div.dat.pos,  # data set to bootstrap       
  statistic = boot_L1L2_pval,  # bootstrap function      
  strata = div.dat.pos$StrataVars, # stratification variable 
  R = Niter)  # number of iterations   

bootres_L1L2_pval_def
bootres_L1L2_pval_pos
```

##### Save the group comparison results

```{r eval=F, echo=T, message=F}
save(bootres_L1L2_pval, file="DPA_results.pval.def.rds")
save(bootres_L1L2_pval, file="DPA_results.pval.pos.rds")
```

##### If we already have saved results, we can load them.

```{r eval=F, echo=T, message=F}
load("DPA_results.pval.def.rds")
load("DPA_results.pval.pos.rds")
```

##### Compute the p-values

```{r eval=F, echo=T, message=F}
pval.def <- round(mean(bootres_L1L2_pval_def$t[,1]>=bootres_L1L2_lmer_def$t0[1], na.rm=T), 5)        
pval.pos <- round(mean(bootres_L1L2_pval_pos$t[,1]>=bootres_L1L2_lmer_pos$t0[1], na.rm=T), 5) 

pval.def
pval.pos
```
